


DEG selection for ingenuity pathway analysis:
p.adj <0.05 with minimal of 500 genes.

1) compound level clustering: use p values, 0 for no pathway, -log10(p) > 1.33 and up for significant pathways
2) dili group level clustering: fraction of compounds where compound is in.

For both cases only use -log10(p) > 1.33  pathays lists

when loading the data make sure to obtain necessary info for both approaches
When myDF is completed enter other script for analysis & visualization





```{r }
# remove points from file names:
options(stringsAsFactors = FALSE)
setwd("H:\\R_HOME\\R_WORK\\mArray BAC reporter studie\\results\\TING pathway analysis\\DEG voor TING\\DEG high 24h BA\\IPA")
#for ( i in seq_along(dir()))
 # {
  #currentPath <- paste(getwd(), folders[i], sep = "/")
  #file.copy( paste(currentPath, dir(currentPath), sep ="/"), gsub(' adj.P_0.05.', 'adjP_005',paste(currentPath, dir(currentPath), sep = "/")))
  #}


rm(list=ls())

library(xlsx)
library(stringr)

rootPath <- "H:\\R_HOME\\R_WORK\\mArray BAC reporter studie\\results\\TING pathway analysis\\DEG voor TING\\DEG high 24h BA\\IPA"
myFolders <- dir(rootPath)
myPaths <- paste(rootPath, myFolders, sep = "\\")
myFile <- dir(myPaths)
pathToFiles <- paste(myPaths, myFile, sep = "\\")
if ( length(unique(myFile)) != length(unique(myFolders))  )
        {
        stop("missing or duplicate files in folders")
        }

testFolderContent <- cbind( myFolders  , myFile )
pat <- "([^/]*_.csv)$"

for ( i in 1 : length( myFolders ) )
  {

  #pat <- "([^ID]*)$"
  pat <- "([^/]*_.xls)$"
      myFile[ i ]
        test<-gsub( pat, "", myFile[ i ])
          #test <- gsub(".ID", "", test)
  test <- gsub("_High_24hr_Control_24hr_rma_log2FCvan1_3 aveExpr_7adjP_005Pathwayanalysis.xls", '', test)
  test <- gsub("_High_24hr_Control_24hr_rma_log2FCvan1_3 aveExpr_7 adjP_005Pathwayanalysis.xls", '', test)
    if ( is.na(test))
      {
      stop(c("check file name in:",  myFolders [ i] ) )
      }
        if ( tolower( test ) != tolower( myFolders[ i ] ) )
          {
          stop ( c("file name not matching folder name at: ", myFolders[i] )  )
          }
        if (nchar(test) <3 )
          {
          stop(myFolders[i])
          }
  }




allPathways = character()
rawList = list( )
myList = list( )
nameList = list()
for ( i in 1 : length( myFolders ) )
  {

  rawList[[ i ]]<- read.xlsx(  pathToFiles[ i ], 1 , startRow = 2, encoding="UTF-8")  
    if(nrow((rawList[[ i ]])) != 0 ){
  myList[[ i ]] <- rawList[[ i ]][ , 1:2 ]
   
  
  
  colnames(myList[[ i ]]) <- c("Ingenuity_Canonical_Pathways", "log10Pvalues")
  } else  {
      rawList[[ i ]] <- data.frame(Ingenuity_Canonical_Pathways = NA, log10Pvalues =NA)
    }
     # myList[[ i ]] <- myList[[i]][ myList[[ i ]][, 2 ] >= 1.33, ]
   
 
nameList[[ i ]] <- myFolders[ i ] 
allPathways <-  c(allPathways, nameList[[ i ]])  #allPathways is variable for all compounds
  if ( i > 1) 
     {

    if(nrow((rawList[[ i-1 ]])) != 1 & nrow((rawList[[ i  ]])) != 1 ){ 
    if ( all( myList[[ i ]][1:20,1] == myList [[ i - 1]][1:20, 1]  ) )
  {
  stop("duplicate file?")
  }
     }
  
  }
}




# gather relevent pathways for Paper Bram

# pull out nfkb and nrf2 pathways for Brams paper
# pathway names: NF-??B Signaling & NRF2-mediated Oxidative Stress Response
# more to test:
#LPS/IL-1 Mediated Inhibition of RXR Function
#Death Receptor Signaling
#TNFR2 Signaling
#Interferon Signaling
#TNFR1 Signaling
#Toll-like Receptor Signaling
NF_kB_Signaling = list()
NRF2_mediated_Oxidative_Stress_Response= list()
LPS_IL_1_Mediated_Inhibition_of_RXR_Function= list()
Death_Receptor_Signaling= list()
TNFR2_Signaling= list()
Interferon_Signaling= list()
TNFR1_Signaling= list()
Toll_like_Receptor_Signaling= list()
for (i in  seq_along(myList)){
  
  
  # store the pathways in data.frame and then later transpose them before rbinding the dataframes in the list
NF_kB_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NF-??B Signaling",]
 if(!is.null(myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NF-??B Signaling",])){
NF_kB_Signaling[[i]][1,1]<- nameList[[i]]
}

NRF2_mediated_Oxidative_Stress_Response[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NRF2-mediated Oxidative Stress Response",]
if(!is.null(NRF2_mediated_Oxidative_Stress_Response[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NRF2-mediated Oxidative Stress Response",])){
NRF2_mediated_Oxidative_Stress_Response[[i]][1,1]<- nameList[[i]]
}
LPS_IL_1_Mediated_Inhibition_of_RXR_Function[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "LPS/IL-1 Mediated Inhibition of RXR Function",]
if(!is.null(LPS_IL_1_Mediated_Inhibition_of_RXR_Function[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "LPS/IL-1 Mediated Inhibition of RXR Function",])){
LPS_IL_1_Mediated_Inhibition_of_RXR_Function[[i]][1,1]<- nameList[[i]]
}

Death_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Death Receptor Signaling",]
if( !is.null(Death_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Death Receptor Signaling",])){
Death_Receptor_Signaling[[i]][1,1]<- nameList[[i]]
}

TNFR2_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR2 Signaling",]
if(!is.null(TNFR2_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR2 Signaling",])){
TNFR2_Signaling[[i]][1,1]<- nameList[[i]]
}

Interferon_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Interferon Signaling",]
if(!is.null(Interferon_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Interferon Signaling",])){
Interferon_Signaling[[i]][1,1]<- nameList[[i]]
}

TNFR1_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR1 Signaling",]
if(!is.null(TNFR1_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR1 Signaling",])){
TNFR1_Signaling[[i]][1,1]<- nameList[[i]]
}

Toll_like_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Toll-like Receptor Signaling",]
if(!is.null(Toll_like_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Toll-like Receptor Signaling",])){
Toll_like_Receptor_Signaling[[i]][1,1]<- nameList[[i]]
}

}


unlist(nameList)

NF_kB_Signaling<-do.call("rbind", NF_kB_Signaling)
colnames(NF_kB_Signaling) <- c("NF_kB_Signaling", paste("NF_kB_Signaling", "log10Pvales", sep = "_"))

NRF2_mediated_Oxidative_Stress_Response<-do.call("rbind", NRF2_mediated_Oxidative_Stress_Response)
colnames(NRF2_mediated_Oxidative_Stress_Response) <- c("NRF2_mediated_Oxidative_Stress_Response", paste("NRF2_mediated_Oxidative_Stress_Response", "log10Pvales", sep = "_"))

LPS_IL_1_Mediated_Inhibition_of_RXR_Function<-do.call("rbind", LPS_IL_1_Mediated_Inhibition_of_RXR_Function)
colnames(LPS_IL_1_Mediated_Inhibition_of_RXR_Function) <- c("LPS_IL_1_Mediated_Inhibition_of_RXR_Function", paste("LPS_IL_1_Mediated_Inhibition_of_RXR_Function", "log10Pvales", sep = "_"))

Death_Receptor_Signaling<-do.call("rbind", Death_Receptor_Signaling)
colnames(Death_Receptor_Signaling) <- c("Death_Receptor_Signaling", paste("Death_Receptor_Signaling", "log10Pvales", sep = "_"))

TNFR2_Signaling<-do.call("rbind", TNFR2_Signaling)
colnames(TNFR2_Signaling) <- c("TNFR2_Signaling", paste("TNFR2_Signaling", "log10Pvales", sep = "_"))

TNFR1_Signaling<-do.call("rbind", TNFR1_Signaling)
colnames(TNFR1_Signaling) <- c("TNFR1_Signaling", paste("TNFR1_Signaling", "log10Pvales", sep = "_"))

Interferon_Signaling<-do.call("rbind", Interferon_Signaling)
colnames(Interferon_Signaling) <- c("Interferon_Signaling", paste("Interferon_Signaling", "log10Pvales", sep = "_"))

Toll_like_Receptor_Signaling<-do.call("rbind", Toll_like_Receptor_Signaling)
colnames(Toll_like_Receptor_Signaling) <- c("Toll_like_Receptor_Signaling", paste("Toll_like_Receptor_Signaling", "log10Pvales", sep = "_"))

#combine all in 1 DF
outputDF <- data.frame(compound = unlist(nameList), row.names = unlist(nameList))

ind <- match( rownames(outputDF),NF_kB_Signaling[,1])
outputDF$NF_kB_Signaling <- NF_kB_Signaling[ind,2]

ind <- match( rownames(outputDF),NRF2_mediated_Oxidative_Stress_Response[,1])
outputDF$NRF2_mediated_Oxidative_Stress_Response <- NRF2_mediated_Oxidative_Stress_Response[ind,2]

ind <- match( rownames(outputDF),LPS_IL_1_Mediated_Inhibition_of_RXR_Function[,1])
outputDF$LPS_IL_1_Mediated_Inhibition_of_RXR_Function <- LPS_IL_1_Mediated_Inhibition_of_RXR_Function[ind,2]

ind <- match( rownames(outputDF),Death_Receptor_Signaling[,1])
outputDF$Death_Receptor_Signaling <- Death_Receptor_Signaling[ind,2]

ind <- match( rownames(outputDF),TNFR2_Signaling[,1])
outputDF$TNFR2_Signaling <- TNFR2_Signaling[ind,2]

ind <- match( rownames(outputDF),TNFR1_Signaling[,1])
outputDF$TNFR1_Signaling <- TNFR1_Signaling[ind,2]

ind <- match( rownames(outputDF),Interferon_Signaling[,1])
outputDF$Interferon_Signaling <- Interferon_Signaling[ind,2]

ind <- match( rownames(outputDF),Toll_like_Receptor_Signaling[,1])
outputDF$Toll_like_Receptor_Signaling <- Toll_like_Receptor_Signaling[ind,2]


outputDF[is.na(outputDF)] <- 0
(outputDF$NRF2_mediated_Oxidative_Stress_Response)[1:10]


require(reshape2)
require(ggplot2)
require(plyr)
head(outputDF)
outputDF$compound <-rownames(outputDF)

outputDF_L <- melt(outputDF, id.vars="compound")
outputDF_L$value <- gsub(",",".", outputDF_L$value)
outputDF_L$value<-as.numeric(outputDF_L$value)
outputDF_L$value[1:10]
head(outputDF_L)
colnames(outputDF_L) <- c("compound","pathway", "min_logpValue")

ggplot(data = outputDF_L, aes(x= compound, y= min_logpValue))  + geom_bar(stat="identity") + facet_wrap( .~pathway)

ddply(outputDF_L, .(compound), summarize, ll=length(compound))


tail(outputDF_L)

myList[[i]]
head(allPathways)

  #1: first pull out all the unique pathways found in myList and create a data.frame
  #2: Then using matching to pull out appropriate values from myList

  rawData <- do.call("rbind", myList)   
  

head(rawData)
dim(rawData)
length(unique(rawData$Ingenuity_Canonical_Pathways))
# create empty data.frame with row names the unique pathways

myDF <- as.data.frame( matrix(NA, nrow = length( unique ( rawData$Ingenuity_Canonical_Pathways )), ncol = length( allPathways ) ) )
  rownames( myDF ) <- unique( rawData$Ingenuity_Canonical_Pathways )
    colnames( myDF ) <- allPathways
      head( myDF )
 


 for ( i in 1 : ncol( myDF )) {
   
   ind <- match( myList[[ i ]]$Ingenuity_Canonical_Pathways,  rownames( myDF ) ) 
   colnames(myList[[ i ]])[2] <- "nlog10Pvalues"
   myDF[ind, i ] <- myList[[ i ]]$nlog10Pvalues 
 
   }

head(myDF)

write.table(file = "IPA_high_8hrs_sel.csv", myDF, sep = ",", col.names = NA)

myDF[is.na( myDF )] <- 0
  head(myDF)

myDFsel <- subset(myDF, apply(as.matrix(myDF), MARGIN = 1, sum  ) > 50  ) #remove some pathways that are not doing much


head(myDFsel)[ , 1:2]
dim(myDFsel)
head(myDF)[, 1:2 ]
dim(myDF)

myDF <- myDFsel

```

bellow loading of compound - annotations - re-do after recreating myDF
```{r }
library(gdata)
options(stringsAsFactors=FALSE)

myGroups <- read.xls( file.path( "/home/winks/compounds for screen",  "TGPDrugsAnnotatedbyLabel_ForSteven.xls"), sep =",", header = T)

myAnot <- data.frame(compounds = colnames(myDF) ) #currently used compounds

myGroups <- subset( myGroups, select = c("Name_TGP","SeverityClass", "DILIConcern" ))
head(myGroups)
colnames(myGroups)<- c("compound", "SeverityClass", "DILIConcern" )   

unique(myGroups$DILIConcern)
myGroups$DILIConcern <- gsub("Less-DILI-concern", "Less-DILI-Concern",   myGroups$DILIConcern )
myGroups$DILIConcern <- gsub("Most-DILI-concern", "Most-DILI-Concern",  myGroups$DILIConcern )
myGroups$DILIConcern[myGroups$DILIConcern == ""] <-"unknown"
#8: fatal hepatoxicity  7: acute liver failure    6: liver necrosis  	5: Jaundice	4: hyperbilirubinemia		3: liver aminotransferases increase		2: cholestasis	1: steatosis

myGroups$SeverityClass[myGroups$SeverityClass == ""] <-"unknown"
myGroups$SeverityClass <-gsub("0", "No_DILI_concern",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("1", "steatosis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("2", "cholestasis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("3", "liver_aminotransferase_increase",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("4", "hyperbilirubinemia",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("5", "jaundice",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("6", "liver_necrosis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("7", "acute_liver_failure",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("8", "fatal_hepatotoxicity",myGroups$SeverityClass   )

# parsing problem for TNF
myGroups$compound[ 167 ] <- "TNF"
myGroups<-myGroups[-171,]


ind<-match(myAnot[,1], myGroups$compound)

if (length(myAnot[,1]) != length(ind)) 
  {
  stop("not guud, some compound did not match in the two lists!")
  }

myAnot$SeverityClass <- myGroups$SeverityClass[ ind ]
myAnot$DILIConcern <- myGroups$DILIConcern[ ind ]

myAnot$SeverityClass <- factor(myAnot$SeverityClass)
myAnot$DILIConcern <- factor(myAnot$DILIConcern)
myAnot$compounds <- factor(myAnot$compounds)


```


```{r}

library(pheatmap)   
library(gplots)
library("Biobase")
library("RColorBrewer")
library(cluster)
library(fpc)
library(ggplot2)
library(kohonen)
library(hopach)
library(vegan)


head(myAnot)
head(myDF)
dim(myDF)

# compound annotations
rownames(myAnot)<- myAnot$compounds
myAnot$compounds <-NULL
head(myAnot)
levels(myAnot$DILIConcern)
levels(myAnot$SeverityClass)
SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
DILIConcern <- c("grey","green", "orange", "darkred")
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
names(DILIConcern) <- c("unknown",  "No-DILI-Concern", "Less-DILI-Concern", "Most-DILI-Concern")
myColors <- list(SeverityClass = SeverityClass, DILIConcern=DILIConcern)

-log10(0.05)
max(myDF)
head(myAnot)
myVector<- cbind(as.character(myAnot$SeverityClass ), colnames(myDF)) 
myVectorInd<- myVector[,1] !="unknown"

myDFtest <- subset(myDF, select = myVectorInd )   #remove non-classified drugs
myDF <- myDFtest


class(myAnot)


myAnottest <- myAnot[ myVectorInd,]
myAnot <- myAnottest

head(myDFtest)
dim(myDF)
dim(myAnot)

#define colors for the pheatmap
## set color representation for specific values of the data distribution
quantile_range <- quantile(sqrt(as.matrix(myDF) ), probs = seq(0, 1, 0.1)) 
as.numeric(quantile_range)
#color_palette <- colorRampPalette(c(greenred(256)[1]  , greenred(256)[128], greenred(256)[256]  ))(length(quantile_range) - 1)

myDF <- sqrt(myDF)
head(myDF)
minV <- min(as.matrix(myDF))
maxV <- max(myDF)



hmcols<-colorRampPalette(c("white","blue"))(256)


white <- hmcols[1]
lightestBlue <-hmcols[30]
lightBlue <- hmcols[80]
blue <- hmcols[128]
darkBlue <- hmcols[156]
darkerBlue <- hmcols[180]
darkestBlue <-hmcols[256]

x<-1:10
hist(x, col = darkestBlue)

sqrt(1.33)
Breakwhite <- sqrt(1.29) 
BreaklightestBlue <- sqrt(2)
BreaklightBlue<- sqrt(2.5)
Breakblue <-  sqrt(3)
BreakdarkBlue <- sqrt( 3.5)
BreakdarkerBlue <- sqrt( 4)
BreakdArkestBlue <- maxV

max(myDF)

  myBreaks <-c(minV, Breakwhite, BreaklightestBlue, BreaklightBlue, Breakblue, BreakdarkBlue, BreakdarkerBlue, BreakdArkestBlue) 


color_palette <- c(white, lightestBlue, lightBlue, blue, darkBlue, darkerBlue, darkestBlue)

the distance measure to be used. This must be one of "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski". Any unambiguous substring can be given. and "correlation" - = pearson

the agglomeration method to be used. This should be (an unambiguous abbreviation of) one of "ward", "single", "complete", "average", "mcquitty", "median" or "centroid".

geneSet <-"IPA"
folderNameSampleSet <- "high_8hrs_selection"
#png(file = paste( geneSet, folderNameSampleSet ,zehDist, zehClust ,".png", sep = "_"), width = 5000, height = 10000, units = "px",  res = 8, type = "Xlib")




# = == this part is to divide myDF in two groups sDILI and nsDILI

head(myDF)
head(myAnot)

all(colnames(myDF )  == rownames(myAnot) )

indS <-  myAnot$DILIConcern %in% "Most-DILI-Concern"
myDF_sDILI <- myDF[, indS]
myAnot_sDILI <- myAnot[ indS, ]

indnS <-  myAnot$DILIConcern %in% "Less-DILI-Concern"
myDF_nsDILI <- myDF[ , indnS ]
myAnot_nsDILI <- myAnot[indnS , ]

head(myAnot_sDILI)

SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
myColors_sDILI <- list(SeverityClass = SeverityClass)



  head(myAnot_nsDILI)

SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
myColors_nsDILI <- list(SeverityClass = SeverityClass)

  
  
# this part is to store clustering of pheatmap into text file
?hclust
?dist
method  
zehDist <- c("manhattan","correlation", "euclidean")[1]
zehClust <- c("average", "ward", "complete", "centroid")[2]
dsDILI <- dist(myDF_sDILI , method = zehDist)
dnDILI <- dist(myDF_nsDILI , method = zehDist)
osDILI <- hclust(dsDILI, method = zehClust)
onsDILI <- hclust(dnDILI, method = zehClust)

TdsDILI <- dist(t(myDF_sDILI) , method = zehDist)
TdnDILI <- dist(t(myDF_nsDILI) , method = zehDist)
TosDILI <- hclust(TdsDILI, method = zehClust)
TonsDILI <- hclust(TdnDILI, method = zehClust)

ordered_myDF_sDILI <- myDF_sDILI[osDILI$order, TosDILI$order  ]
ordered_myDF_nsDILI <- myDF_nsDILI[onsDILI$order, TonsDILI$order  ]


# also try kmeans
kmeans2 <- kmeans(t(myDF_sDILI), center =3,  nstart =25)
kmeans2$cluster
o <- order(kmeans2$cluster)
m2 <- (t(myDF_sDILI)[o , ])
#genes
kmeans2G <- kmeans(myDF_sDILI, center =5,  nstart =25)
oG <- order(kmeans2G$cluster)
sDILIkmeans <- (m2[ , oG ])



kmeans2 <- kmeans(t(myDF_nsDILI), center =3,  nstart =25)
kmeans2$cluster
o <- order(kmeans2$cluster)
m2 <- (t(myDF_nsDILI)[o , ])
#genes
kmeans2G <- kmeans(myDF_nsDILI, center =3,  nstart =25)
oG <- order(kmeans2G$cluster)
nsDILIkmeans <- (m2[ , oG ])



pdf(file = paste("nsDILIkmeans_3_3", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
pheatmap(as.matrix(t(nsDILIkmeans)), col= color_palette, breaks =  myBreaks , main= paste(folderNameSampleSet,geneSet) ,  
         fontsize_row=2, fontsize_col =6, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
          legend = TRUE, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()


# now reorder rows to match each other and then recluster only the rows based on combined data

ind <- match( rownames(t(nsDILIkmeans) ), rownames(t(sDILIkmeans ) ) )
reordered_sDILIkmeans <- t(sDILIkmeans)[ind, ]
all(rownames(reordered_sDILIkmeans) == rownames(t(nsDILIkmeans) ) )
allData <- cbind(reordered_sDILIkmeans, t(nsDILIkmeans) )
head(allData)
allData<-as.data.frame(allData)


indBoth <-  myAnot$DILIConcern %in% c("Most-DILI-Concern", "Less-DILI-Concern")

myAnot_bothDILI <- myAnot[ indBoth, ]


SeverityClass <- c(  "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
DILIConcern <- c( "orange", "darkred")
names(SeverityClass) <- c(  "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
names(DILIConcern) <- c("Less-DILI-Concern", "Most-DILI-Concern")
myColors <- list(SeverityClass = SeverityClass, DILIConcern=DILIConcern)

zehDist <- c("manhattan","correlation", "euclidean")[1]
zehClust <- c("single" , "average", "ward", "complete", "centroid")[2]

pdf(file = paste("BothSepKmeans", geneSet, folderNameSampleSet ,zehDist, zehClust ,".pdf", sep = "_"), width = 8, height = 10, pointsize = 1)
fontsize_row = 3; fontsize_col = 3;
pheatmap(as.matrix(allData), col= color_palette, breaks =  myBreaks , main= paste(geneSet, folderNameSampleSet, sep ="_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = TRUE, cluster_cols = FALSE, clustering_distance_rows = zehDist, 
         clustering_method = zehClust,  
          legend = TRUE, fontsize = 4,
          legend_breaks = round(myBreaks, digits=2), 
         annotation = myAnot_bothDILI, annotation_colors = myColors,
         
         display_numbers = F )
dev.off()







#end of the splitting thing for Bob

zehDist <- c("manhattan","correlation", "euclidean")[3]
zehClust <- c("average", "ward", "complete", "centroid")[4]
pdf(file = paste( geneSet, folderNameSampleSet ,zehDist, zehClust ,"SEL.pdf", sep = "_"), width = 8, height = 10, pointsize = 1)
fontsize_row = 3; fontsize_col = 3;
pheatmap(as.matrix(myDF), col= color_palette, breaks =  myBreaks , main= paste(geneSet, folderNameSampleSet, sep ="_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = TRUE, cluster_cols = TRUE, clustering_distance_rows = zehDist, 
         clustering_distance_cols = zehDist, clustering_method = zehClust,  
          legend = TRUE, fontsize = 4,
          legend_breaks = round(myBreaks, digits=2), 
         annotation = myAnot, annotation_colors = myColors,
         
         display_numbers = F )
dev.off()

  ?pheatmap
#PCA plots: for complete code for pca plot including explaining variables (so genes) see : http://stackoverflow.com/questions/6578355/plotting-pca-biplot-with-ggplot2
?prcomp

myDFT<-t(myDF)

head(myDFT)
prin_comp<-rda(myDFT, scale=TRUE)
pca_scores<-scores(prin_comp)
head(pca_scores)

pdf(file = paste("PCA", folderNameSampleSet, ".pdf", sep  ="_"), height = 6, width = 6 )
plot(pca_scores$sites[,1],
     pca_scores$sites[,2],
     pch=21,
     bg=as.numeric(factor( myAnot$DILIConcern ) ),
     xlim=c(-2,2), 
     ylim=c(-2,2), 
     main = paste(geneSet,folderNameSampleSet) )
ordiellipse(prin_comp,myAnot$DILIConcern,conf=0.8)     
legend(x=1.3, y = 2, legend = c("unknown (blue)", "sDILI (red)", "nsDILI (black)", "nonDili (green)"  ), cex = 0.6)     
dev.off()

?legend
head(myAnot$DILIConcern)
as.numeric(myAnot$DILIConcern)
?plot  
    

To make ellipses, function ordiellipse() of package vegan is used. As arguments PCA analysis object and grouping variable must be provided. To control number of points included in ellipse, argument conf= can be used.

x<-1:10
hist(x, col = 2)


# clustering: how many clusters? take care if it is sample or gene based


#"euclidean" -2 , "maximum" -2, "manhattan" -5 , "canberra" -2, "binary" or "minkowski -2"
myDistDF <- dist(myDFT, method = "manhattan") # sample clustering
myDistDFM <- (as.matrix(myDistDF))
silEst <- silcheck(myDistDFM, diss = TRUE)
silEst

#"euclidean" -2 , "maximum" -2, "manhattan" -2 , "canberra" -2, "binary" or "minkowski -2"
myDistDF <- dist(myDF, method = "euclidean") # gene clustering
myDistDFM <- (as.matrix(myDistDF))
silEst <- silcheck(myDistDFM, diss = TRUE)
silEst

hobach.hobj <- hopach(myDFT, dmat = myDistDF)
hobach.hobj$clust$k 
hobach.hobj$cluster$sizes 

#conclusion: 2 or 4 gene clusters and 2,4,5 or 1 sample clusters

# partition based clustering: kmeans


#samples
kmeans2 <- kmeans(myDFT, center =3,  nstart =25)
kmeans2$cluster

kmeansoutput <- cbind(kmeans2$cluster )
o <- order(kmeans2$cluster)
m2 <- (myDFT[o , ])
#genes
kmeans2G <- kmeans(myDF, center =3,  nstart =25)

oG <- order(kmeans2G$cluster)
m2G <- (m2[ , oG ])


pdf(file = paste("kmeans_3_3", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
pheatmap(as.matrix(t(m2G)), col= color_palette, breaks =  myBreaks , main= paste(folderNameSampleSet,geneSet) ,  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
          legend = TRUE, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()




#8: fatal hepatoxicity  7: acute liver failure    6: liver necrosis    5: Jaundice	4: hyperbilirubinemia		3: liver aminotransferases increase		2: cholestasis	1: steatosis
    
    
    
write.table(kmeansoutput, file = paste(folderNameSampleSet, "kmeansC2", "csv", sep = "." ), sep = "," )



kmeans2 <- kmeans(myDFT, center =2,  nstart =25)
pdf(file = paste("Kmeans_clusterPlot", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
clusplot(myDFT, kmeans2$cluster, color= TRUE , shade=TRUE, 
         labels=2, lines=0, cex = 0.3, main = paste("clusterPlot", folderNameSampleSet, geneSet,  sep = "_") )
dev.off()

?clusplot
# partition based clustering: PAM

# euclidian and manhatten
pamK2.myDF <- pam(myDFT, diss = FALSE, metric= "manhattan", k = 2)
pamK2G.myDF <- pam(myDF, diss = FALSE, metric= "manhattan", k = 2)

o<-order(pamK2.myDF$cluster)
oG<-order(pamK2G.myDF$cluster)
pamM2 <- myDFT[o, oG  ]


pdf(file = paste("PAM_2_2_manh", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)

pheatmap(as.matrix(t(pamM2)), col= color_palette, breaks =  myBreaks , main= paste("PAM_S2_G3", folderNameSampleSet, geneSet,  sep = "_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE,  fontsize = 6,
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()




    pamResults <- cbind(pamK2.myDF$cluster)
write.table(pamResults, file = paste(folderNameSampleSet, "PAM_k5", "csv", sep = "." ), sep = "," )

pamK2.myDF <- pam(myDFT, diss = FALSE, metric= "euclidean", k = 2)
pdf(file = paste("PAMK2_clusterplot_eucl", folderNameSampleSet, geneSet, ",pdf", sep = "_") , width = 8, height = 8)
clusplot(myDFT, pamK2.myDF$cluster, color= TRUE , shade=TRUE, 
         labels=2, lines=0, cex = 0.3, main = paste("PAMK2_clusterPlot", folderNameSampleSet, geneSet,  sep = "_") )
dev.off()



# partition based clustering: SOM



SOM.myDF <- som(myDFT, grid = somgrid( 3, 3))
SOMG.myDF <- som(as.matrix(myDF), grid = somgrid( 2, 2 ))

table(SOM.myDF$unit.classif)

o<-order(SOM.myDF$unit.classif)
oG<-order(SOMG.myDF$unit.classif)
SOMM2 <- myDFT[o, oG  ]
head(t(SOMM2))


pdf(file = paste( "SOM_3_2", folderNameSampleSet, geneSet, ".pdf", sep = "_" ), width = 8, height = 10)

pheatmap(as.matrix(t(SOMM2)), col= color_palette, breaks =  myBreaks , main=paste( "SOM_S3_G4", folderNameSampleSet, geneSet, sep = "_" ),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE, display_numbers = F, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()

        cosangle
euclid
?distancematrix
myDist <- distancematrix(myDFT, d = "cosangle")
hobj <- hopach( myDFT, dmat = myDist)
myDist <- distancematrix(myDF, d = "cosangle")
hobjG <- hopach( myDF, dmat = myDist)

table(hobj$clustering$labels)
table(hobjG$clustering$labels)

o<-order(hobj$clustering$labels)
oG<-order(hobjG$clustering$labels)
HOP <- myDFT[o, oG  ]
head(t(HOP))


pdf(file = paste("HM_hopach_cosangle", folderNameSampleSet, geneSet,".pdf", sep = "_")  , width = 8, height = 10 )

pheatmap(as.matrix(t(HOP ) ), col= color_palette, breaks =  myBreaks , main= paste("HM_hopach_euclid", folderNameSampleSet, geneSet, sep = "_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE, display_numbers = F, fontsize = 6,
          legend_breaks = round(myBreaks, digits = 2), treeheight_row =0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()





  HOB1 <-cbind(rownames(myDFT[hobj$clust$labels == 300, ]), rep(1, sum(hobj$clust$labels == 300 )  ))
HOB2 <-cbind(rownames(myDFT[hobj$clust$labels == 250, ]), rep(2, sum(hobj$clust$labels == 250 )  ))

hopachgrps <- rbind(HOB1, HOB2)
write.table(hopachgrps, file = paste(folderNameSampleSet, "Hopach_2_biggest", "csv", sep = "." ), sep = "," )







```



