


DEG selection for ingenuity pathway analysis:
p.adj <0.05 with minimal of 500 genes.

1) compound level clustering: use p values, 0 for no pathway, -log10(p) > 1.33 and up for significant pathways
2) dili group level clustering: fraction of compounds where compound is in.

For both cases only use -log10(p) > 1.33  pathays lists

when loading the data make sure to obtain necessary info for both approaches
When myDF is completed enter other script for analysis & visualization





```{r }
# remove points from file names:
options(stringsAsFactors = FALSE)
options(encoding ="UTF-8")
setwd("H:\\R_HOME\\R_WORK\\mArray BAC reporter studie\\results\\TING pathway analysis\\DEG voor TING\\DEG high 24h BA\\IPA")
rm(list=ls())
#for ( i in seq_along(dir()))
 # {
  #currentPath <- paste(getwd(), folders[i], sep = "/")
  #file.copy( paste(currentPath, dir(currentPath), sep ="/"), gsub(' adj.P_0.05.', 'adjP_005',paste(currentPath, dir(currentPath), sep = "/")))
  #}


rm(list=ls())

library(xlsx)
library(stringr)

rootPath <- "H:\\R_HOME\\R_WORK\\mArray BAC reporter studie\\results\\TING pathway analysis\\DEG voor TING\\DEG high 24h BA\\IPA"
myFolders <- dir(rootPath)
myPaths <- paste(rootPath, myFolders, sep = "\\")
myFile <- dir(myPaths)
pathToFiles <- paste(myPaths, myFile, sep = "\\")
if ( length(unique(myFile)) != length(unique(myFolders))  )
        {
        stop("missing or duplicate files in folders")
        }

testFolderContent <- cbind( myFolders  , myFile )
pat <- "([^/]*_.csv)$"

for ( i in 1 : length( myFolders ) )
  {

  #pat <- "([^ID]*)$"
  pat <- "([^/]*_.xls)$"
      myFile[ i ]
        test<-gsub( pat, "", myFile[ i ])
          #test <- gsub(".ID", "", test)
  test <- gsub("_High_24hr_Control_24hr_rma_log2FCvan1_3 aveExpr_7adjP_005Pathwayanalysis.xls", '', test)
  test <- gsub("_High_24hr_Control_24hr_rma_log2FCvan1_3 aveExpr_7 adjP_005Pathwayanalysis.xls", '', test)
    if ( is.na(test))
      {
      stop(c("check file name in:",  myFolders [ i] ) )
      }
        if ( tolower( test ) != tolower( myFolders[ i ] ) )
          {
          stop ( c("file name not matching folder name at: ", myFolders[i] )  )
          }
        if (nchar(test) <3 )
          {
          stop(myFolders[i])
          }
  }




allPathways = character()
rawList = list( )
myList = list( )
nameList = list()
for ( i in 1 : length( myFolders ) )
  {
  rawList[[ i ]]<- read.xlsx(  pathToFiles[ i ], 1 , startRow = 2, encoding="UTF-8")  
    if(nrow((rawList[[ i ]])) != 0 ){
  myList[[ i ]] <- rawList[[ i ]][ , 1:2 ]
   
  
  
  colnames(myList[[ i ]]) <- c("Ingenuity_Canonical_Pathways", "log10Pvalues")
  } else  {
      rawList[[ i ]] <- data.frame(Ingenuity_Canonical_Pathways = NA, log10Pvalues =NA)
    }
     # myList[[ i ]] <- myList[[i]][ myList[[ i ]][, 2 ] >= 1.33, ]
   
 
nameList[[ i ]] <- myFolders[ i ] 
allPathways <-  c(allPathways, nameList[[ i ]])  #allPathways is variable for all compounds
  if ( i > 1) 
     {

    if(nrow((rawList[[ i-1 ]])) != 1 & nrow((rawList[[ i  ]])) != 1 ){ 
    if ( all( myList[[ i ]][1:20,1] == myList [[ i - 1]][1:20, 1]  ) )
  {
  stop("duplicate file?")
  }
     }
  
  }
}


for (kk in seq_along(myList)) {
print(head(myList[[kk]]))
}
# gather relevent pathways for Paper Bram

# pull out nfkb and nrf2 pathways for Brams paper
# pathway names: NF-??B Signaling & NRF2-mediated Oxidative Stress Response
# more to test:
#LPS/IL-1 Mediated Inhibition of RXR Function
#Death Receptor Signaling
#TNFR2 Signaling
#Interferon Signaling
#TNFR1 Signaling
#Toll-like Receptor Signaling
NF_kB_Signaling = list()
NRF2_mediated_Oxidative_Stress_Response= list()
LPS_IL_1_Mediated_Inhibition_of_RXR_Function= list()
Death_Receptor_Signaling= list()
TNFR2_Signaling= list()
Interferon_Signaling= list()
TNFR1_Signaling= list()
Toll_like_Receptor_Signaling= list()
for (i in  seq_along(myList)){
write.table(myList[[i]][grepl("NF-",myList[[i]][,1]),], file = "enc.txt")
  getwd()
  # store the pathways in data.frame and then later transpose them before rbinding the dataframes in the list
NF_kB_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NF-??B Signaling",]
 if(!is.null(myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NF-??B Signaling",])){
NF_kB_Signaling[[i]][1,1]<- nameList[[i]]
}

NRF2_mediated_Oxidative_Stress_Response[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NRF2-mediated Oxidative Stress Response",]
if(!is.null(NRF2_mediated_Oxidative_Stress_Response[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "NRF2-mediated Oxidative Stress Response",])){
NRF2_mediated_Oxidative_Stress_Response[[i]][1,1]<- nameList[[i]]
}
LPS_IL_1_Mediated_Inhibition_of_RXR_Function[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "LPS/IL-1 Mediated Inhibition of RXR Function",]
if(!is.null(LPS_IL_1_Mediated_Inhibition_of_RXR_Function[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "LPS/IL-1 Mediated Inhibition of RXR Function",])){
LPS_IL_1_Mediated_Inhibition_of_RXR_Function[[i]][1,1]<- nameList[[i]]
}

Death_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Death Receptor Signaling",]
if( !is.null(Death_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Death Receptor Signaling",])){
Death_Receptor_Signaling[[i]][1,1]<- nameList[[i]]
}

TNFR2_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR2 Signaling",]
if(!is.null(TNFR2_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR2 Signaling",])){
TNFR2_Signaling[[i]][1,1]<- nameList[[i]]
}

Interferon_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Interferon Signaling",]
if(!is.null(Interferon_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Interferon Signaling",])){
Interferon_Signaling[[i]][1,1]<- nameList[[i]]
}

TNFR1_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR1 Signaling",]
if(!is.null(TNFR1_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "TNFR1 Signaling",])){
TNFR1_Signaling[[i]][1,1]<- nameList[[i]]
}

Toll_like_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Toll-like Receptor Signaling",]
if(!is.null(Toll_like_Receptor_Signaling[[i]] <-  myList[[i]][ myList[[i]]$Ingenuity_Canonical_Pathways== "Toll-like Receptor Signaling",])){
Toll_like_Receptor_Signaling[[i]][1,1]<- nameList[[i]]
}

}


unlist(nameList)

NF_kB_Signaling<-do.call("rbind", NF_kB_Signaling)
colnames(NF_kB_Signaling) <- c("NF_kB_Signaling", paste("NF_kB_Signaling", "log10Pvales", sep = "_"))

NRF2_mediated_Oxidative_Stress_Response<-do.call("rbind", NRF2_mediated_Oxidative_Stress_Response)
colnames(NRF2_mediated_Oxidative_Stress_Response) <- c("NRF2_mediated_Oxidative_Stress_Response", paste("NRF2_mediated_Oxidative_Stress_Response", "log10Pvales", sep = "_"))

LPS_IL_1_Mediated_Inhibition_of_RXR_Function<-do.call("rbind", LPS_IL_1_Mediated_Inhibition_of_RXR_Function)
colnames(LPS_IL_1_Mediated_Inhibition_of_RXR_Function) <- c("LPS_IL_1_Mediated_Inhibition_of_RXR_Function", paste("LPS_IL_1_Mediated_Inhibition_of_RXR_Function", "log10Pvales", sep = "_"))

Death_Receptor_Signaling<-do.call("rbind", Death_Receptor_Signaling)
colnames(Death_Receptor_Signaling) <- c("Death_Receptor_Signaling", paste("Death_Receptor_Signaling", "log10Pvales", sep = "_"))

TNFR2_Signaling<-do.call("rbind", TNFR2_Signaling)
colnames(TNFR2_Signaling) <- c("TNFR2_Signaling", paste("TNFR2_Signaling", "log10Pvales", sep = "_"))

TNFR1_Signaling<-do.call("rbind", TNFR1_Signaling)
colnames(TNFR1_Signaling) <- c("TNFR1_Signaling", paste("TNFR1_Signaling", "log10Pvales", sep = "_"))

Interferon_Signaling<-do.call("rbind", Interferon_Signaling)
colnames(Interferon_Signaling) <- c("Interferon_Signaling", paste("Interferon_Signaling", "log10Pvales", sep = "_"))

Toll_like_Receptor_Signaling<-do.call("rbind", Toll_like_Receptor_Signaling)
colnames(Toll_like_Receptor_Signaling) <- c("Toll_like_Receptor_Signaling", paste("Toll_like_Receptor_Signaling", "log10Pvales", sep = "_"))

#combine all in 1 DF
outputDF <- data.frame(compound = unlist(nameList), row.names = unlist(nameList))

ind <- match( rownames(outputDF),NF_kB_Signaling[,1])
outputDF$NF_kB_Signaling <- NF_kB_Signaling[ind,2]

ind <- match( rownames(outputDF),NRF2_mediated_Oxidative_Stress_Response[,1])
outputDF$NRF2_mediated_Oxidative_Stress_Response <- NRF2_mediated_Oxidative_Stress_Response[ind,2]

ind <- match( rownames(outputDF),LPS_IL_1_Mediated_Inhibition_of_RXR_Function[,1])
outputDF$LPS_IL_1_Mediated_Inhibition_of_RXR_Function <- LPS_IL_1_Mediated_Inhibition_of_RXR_Function[ind,2]

ind <- match( rownames(outputDF),Death_Receptor_Signaling[,1])
outputDF$Death_Receptor_Signaling <- Death_Receptor_Signaling[ind,2]

ind <- match( rownames(outputDF),TNFR2_Signaling[,1])
outputDF$TNFR2_Signaling <- TNFR2_Signaling[ind,2]

ind <- match( rownames(outputDF),TNFR1_Signaling[,1])
outputDF$TNFR1_Signaling <- TNFR1_Signaling[ind,2]

ind <- match( rownames(outputDF),Interferon_Signaling[,1])
outputDF$Interferon_Signaling <- Interferon_Signaling[ind,2]

ind <- match( rownames(outputDF),Toll_like_Receptor_Signaling[,1])
outputDF$Toll_like_Receptor_Signaling <- Toll_like_Receptor_Signaling[ind,2]


outputDF[is.na(outputDF)] <- 0
(outputDF$NRF2_mediated_Oxidative_Stress_Response)[1:10]

require(reshape2)
require(ggplot2)
require(plyr)
head(outputDF)
outputDF$compound <-rownames(outputDF)

outputDF_L <- melt(outputDF, id.vars="compound")
outputDF_L$value <- gsub(",",".", outputDF_L$value)
outputDF_L$value<-as.numeric(outputDF_L$value)
outputDF_L$value[1:10]
(outputDF_L)[140:240,]
colnames(outputDF_L) <- c("compound","pathway", "min_logpValue")







head(outputDF_L)
unique(outputDF_L$compound)



# for bram now move on to annotating in next block


ddply(outputDF_L, .(pathway), summarize, ll=length(pathway))

getwd()
tail(outputDF_L)

myList[[i]]
head(allPathways)

  #1: first pull out all the unique pathways found in myList and create a data.frame
  #2: Then using matching to pull out appropriate values from myList

  rawData <- do.call("rbind", myList)   
  rawData)



head(rawData)
dim(rawData)
length(unique(rawData$Ingenuity_Canonical_Pathways))
# create empty data.frame with row names the unique pathways

myDF <- as.data.frame( matrix(NA, nrow = length( unique ( rawData$Ingenuity_Canonical_Pathways )), ncol = length( allPathways ) ) )
  rownames( myDF ) <- unique( rawData$Ingenuity_Canonical_Pathways )
    colnames( myDF ) <- allPathways
      head( myDF )
# end Bram stuff (right?) 


 for ( i in 1 : ncol( myDF )) {
   
   ind <- match( myList[[ i ]]$Ingenuity_Canonical_Pathways,  rownames( myDF ) ) 
   #colnames(myList[[ i ]])[2] <- "nlog10Pvalues"
   myDF[ind, i ] <- myList[[ i ]]$nlog10Pvalues 
 
   }

head(myDF)

write.table(file = "IPA_high_8hrs_sel.csv", myDF, sep = ",", col.names = NA)

myDF[is.na( myDF )] <- 0
  head(myDF)

myDFsel <- subset(myDF, apply(as.matrix(myDF), MARGIN = 1, sum  ) > 50  ) #remove some pathways that are not doing much


head(myDFsel)[ , 1:2]
dim(myDFsel)
head(myDF)[, 1:2 ]
dim(myDF)

myDF <- myDFsel

```

bellow loading of compound - annotations - re-do after recreating myDF
```{r }
myDF <- outputDF_L
options(stringsAsFactors=FALSE)

myGroups <- read.delim( file.path( "H:\\R_HOME\\R_WORK\\mArray BAC reporter studie",  "TGPDrugsAnnotatedbyLabel_updated082913-1.csv"), sep =",", header = T)

myAnot <- data.frame(compounds = unique(myDF$compound) ) #currently used compounds

myGroups
myGroups <- subset( myGroups, select = c("Name_TGP","SeverityClass", "DILIConcern" ))
head(myGroups)
colnames(myGroups)<- c("compound", "SeverityClass", "DILIConcern" )   

unique(myGroups$DILIConcern)
myGroups$DILIConcern <- gsub("Less-DILI-concern", "Less-DILI-Concern",   myGroups$DILIConcern )
myGroups$DILIConcern <- gsub("Most-DILI-concern", "Most-DILI-Concern",  myGroups$DILIConcern )
myGroups$DILIConcern[myGroups$DILIConcern == ""] <-"unknown"
#8: fatal hepatoxicity  7: acute liver failure    6: liver necrosis  	5: Jaundice	4: hyperbilirubinemia		3: liver aminotransferases increase		2: cholestasis	1: steatosis

myGroups$SeverityClass[myGroups$SeverityClass == ""] <-"unknown"
myGroups$SeverityClass <-gsub("0", "No_DILI_concern",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("1", "steatosis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("2", "cholestasis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("3", "liver_aminotransferase_increase",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("4", "hyperbilirubinemia",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("5", "jaundice",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("6", "liver_necrosis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("7", "acute_liver_failure",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("8", "fatal_hepatotoxicity",myGroups$SeverityClass   )
myGroups$SeverityClass<-factor(myGroups$SeverityClass)
# parsing problem for TNF
#myGroups$compound[ 167 ] <- "TNF"
#myGroups<-myGroups[-171,]

length(unique(myGroups$compound))
ind<-match(myAnot[,1], myGroups$compound)

length(unique(ind))

if (length(myAnot[,1]) != length(ind)) 
  {
  stop("not guud, some compound did not match in the two lists!")
  }

myAnot$SeverityClass <- myGroups$SeverityClass[ ind ]
myAnot$DILIConcern <- myGroups$DILIConcern[ ind ]

myAnot$SeverityClass <- factor(myAnot$SeverityClass)
myAnot$DILIConcern <- factor(myAnot$DILIConcern)
myAnot$compounds <- factor(myAnot$compounds)
head(myDF)
myDFanot <- merge(myDF, myAnot, by.x= "compound", by.y="compounds", all = TRUE)


myDFanot <- myDFanot[ myDFanot$SeverityClass !="unknown",]
myDFanot <- myDFanot[ myDFanot$SeverityClass !="No_DILI_concern",]
myDFanot <- myDFanot[ myDFanot$SeverityClass !="w.d.",]
head(myDFanot)
myDFanot <- within(myDFanot, {
pathway <- reorder(pathway, min_logpValue)
compound <- reorder(compound, -min_logpValue)
})




# select relevant pathways and calculate mean
ox.stress<- myDFanot[ myDFanot$pathway == "NRF2_mediated_Oxidative_Stress_Response", ]
inflammatory <- myDFanot[ myDFanot$pathway == "Toll_like_Receptor_Signaling" | 
                            myDFanot$pathway == "Death_Receptor_Signaling" | 
                            myDFanot$pathway == "TNFR2_Signaling" |
                            myDFanot$pathway == "NF_kB_Signaling" |
                           myDFanot$pathway == "TNFR1_Signaling", ]
head(inflammatory)

inflammatorySUm <- ddply(inflammatory, .(compound,SeverityClass,DILIConcern), function(x) x$meanPval = mean(x$min_logpValue)  )
colnames(inflammatorySUm)[ncol(inflammatorySUm)] <- "meanPval"
head(inflammatorySUm)

unique(myDFanot$pathway)
inflammatorySUm<-inflammatorySUm[ inflammatorySUm$compound != "doxorubicin",]
ox.stress <- ox.stress[ ox.stress$compound != "doxorubicin", ]



source("H:/R_HOME/R_WORK/R scipts/theme_sharp.R")
require(ggplot2)
require(grid)
head(ox.stress)



inflammatorySUm <- within(inflammatorySUm, {
compound <- reorder(compound, -meanPval)
#SeverityClass <- reorder(SeverityClass, -min_logpValue)
})

p<-ggplot(data = inflammatorySUm, aes(x= compound, y= meanPval,  fill = SeverityClass))  + geom_bar(stat="identity") 

p <- p + theme_sharp() + theme( axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.2, size = 12 , colour = "grey50") ) + 
theme( strip.text.x = element_text( size = 12  )) +
  ggtitle( "IPA" ) + 
  theme(plot.title = element_text(lineheight=.8, size = 8 )) 

p
ggsave(file = "inflammatorySUm same legend color.pdf", height = 8, width =14)

save(p,file="inflammatorym stress.RData")
load("inflammatorym stress.RData")
p
#setwd("C:/Users/winks/Documents/own papers/NFKB Nrf2 Bram/figures")




write.table(ox.stress, file = "C:/Users/winks/Documents/own papers/NFKB Nrf2 Bram/figures/oxidativeStress.csv" ,sep =",", col.names = NA)



```


Mijn analyze was als volgt:
Alle stappen 1 tm 5 alleen en enkel op 24hr high dose gedaan.
1)  Oxidatieve stress/ Nrf2 en inflammatory/p65 signaling pathways opgezocht uit: KEGG, REACTOME, GO, prof. TRANSFAC, BIOCARTA & IPA. 
2)	Genen eruit getrokken, heb nu dus een Nrf2 lijstje genen en een inflammatory/ p65 lijstje genen.
3)	Nrf2 lijstje moest significant, redelijke FC (richting maakte niet uit) en minimale gem. expressie hebben voor DEM ofwel BHA (55 genen), anders eruitgeknikkert. Zelfde voor p65 lijstje met IL-1, TNF, LPS(~150 genen). Zelfde selectiecriteria voor bijde lijstjes. 
4)	Stap 3 herhaald maar dan strengere FC selectie voor p65 lijstje, nu 54 genen - dus ongeveer even groot als Nrf2 lijstje.
5)	Enkele heatmaps gemaakt van 24 uur high dose data met de 2 gen lijstjes.
6)	Wil dit nu gaan doen voor 8 uur tenzij die "heatmap 24hrs both 55 genes pearson average.pdf" goed genoeg is.
7)	Dan nog per dili-severity groep appart clusteren- maar dan kunnen alleen de compounds clusteren per groep en niet de genen, zo zou duidelijker moeten worden welke set genen anders in de verschillende dili-groepen tot expressie komen. De genen cluster ik dan over alle dili-groepen tezamen.


Bob: selectie gebaseerd op ook 8 uur data. en heatmap voor controle compounds 8 en 24 uur. Wel 2 apparte heatmaps voor de 2 tijdpunten

```{r}

library(pheatmap)   
library(gplots)
library("Biobase")
library("RColorBrewer")
library(cluster)
#library(fpc)
library(ggplot2)
#library(kohonen)
#library(hopach)
#library(vegan)

# load oxidative stress and inflammatory signaling
# filter based on PHH model compounds: LPS, TNF, IL-1 / DEM/ BHA that is either significant at 8hr or 24hr high dose.
# create heatmap with annotations of DILI severity
# create heatmap with certain DILI-annotations pulled out
rm(list=ls())

inflamData <- read.delim(file = "C:/Users/winks/Documents/own papers/NFKB Nrf2 Bram/NFkB and Nrf2 pathway annotation exports/PHH output/inflammatory.csv",
                         sep =",", header = TRUE)

inflamData<- inflamData[ !inflamData$treatment %in% "doxorubicin",]

oxiData <- read.delim(file = "C:/Users/winks/Documents/own papers/NFKB Nrf2 Bram/NFkB and Nrf2 pathway annotation exports/PHH output/PHH oxidative stress.csv",
                         sep =",", header = TRUE)

  sort(unique(oxiData$treatment))
sum(oxiData$treatment == "doxorubicin")
oxiData <- oxiData[ !oxiData$treatment %in% "doxorubicin", ]
head(oxiData)
head(inflamData)

# select genes for oxiData: model compounds are: diethyl maleate, butylated hydroxyanisole
unique(oxiData$treatment)
subSetOxi <- oxiData[ ( oxiData$treatment == "diethyl maleate" | 
                          oxiData$treatment == "butylated hydroxyanisole" ) & 
                          (oxiData$time == "8hr" | oxiData$time =="24hr" ) &
                      oxiData$dose == "High", ]
         head(subSetOxi)
selOxiGenes <- subSetOxi[ subSetOxi$adj.P.Val <= 0.05 & subSetOxi$AveExpr >=5 & abs(subSetOxi$logFC) > log2(1.5),'ENTREZID']
selOxiGenes <- unique(selOxiGenes)
length(selOxiGenes)


# same for inflammatory: model compounds are: TNF, LPS, interleukin 1 beta, human
unique(inflamData$treatment)
subSetImflam <- inflamData[ ( inflamData$treatment == "TNF" | 
                          inflamData$treatment == "LPS" |
                          inflamData$treatment == "interleukin 1 beta, human" ) & 
                          (inflamData$time == "8hr" | inflamData$time =="24hr" ) &
                      inflamData$dose == "High", ]
         


selInflamGenes <- subSetImflam[ subSetImflam$adj.P.Val <= 0.05 & subSetImflam$AveExpr >=5 & abs(subSetImflam$logFC) > log2(1.5) ,'ENTREZID']
selInflamGenes <- unique(selInflamGenes)
length(selInflamGenes)

inflamData$gene_type <- "Inflammatory"
oxiData$gene_type <- "Oxidative"

myData <- rbind(inflamData, oxiData)
head(myData)
myData.8h <- myData[ (myData$dose == "High" & 
                       myData$time == "8hr") | 
                       (myData$time == "24hr" &
                          myData$dose == "High" & 
                          (myData$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"))) , ]

ind <- myData.8h$ENTREZID %in% unique(c(selOxiGenes, selInflamGenes))
myData.8h.sel <- myData.8h[ind,]

myData.24h <- myData[ (myData$dose == "High" & 
                       myData$time == "24hr") | 
                        (myData$dose == "High" &
                        myData$time == "8hr" & 
                           (myData$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"))) , ]


ind <- myData.24h$ENTREZID %in% unique(c(selOxiGenes, selInflamGenes))
myData.24h.sel <- myData.24h[ind,]

# control compounds for 8hr and 2hr need separate identifier

unique(myData.8h.sel$treatment)


myData.8h.sel$treatment <- as.character(myData.8h.sel$treatment)
myData.8h.sel$time <- as.character(myData.8h.sel$time)

myData.8h.sel[ myData.8h.sel$time == "24hr" & myData.8h.sel$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"), "treatment" ] <-
  
  paste(myData.8h.sel[ myData.8h.sel$time == "24hr" & myData.8h.sel$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"), "treatment" ],
myData.8h.sel[ myData.8h.sel$time == "24hr" & myData.8h.sel$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"), "time" ], sep = "_")

myData.24h.sel

myData.24h.sel$treatment <- as.character(myData.24h.sel$treatment)
myData.24h.sel$time <- as.character(myData.24h.sel$time)

myData.24h.sel[ myData.24h.sel$time == "8hr" & myData.24h.sel$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"), "treatment" ] <-
  
  paste(myData.24h.sel[ myData.24h.sel$time == "8hr" & myData.24h.sel$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"), "treatment" ],
myData.24h.sel[ myData.24h.sel$time == "8hr" & myData.24h.sel$treatment %in% c("TNF", "LPS", "interleukin 1 beta, human","diethyl maleate", "butylated hydroxyanisole"), "time" ], sep = "_")

#oh.. only fda-DILI-annotated compounds

myData.8h.sel <- myData.8h.sel [ myData.8h.sel$DILIConcern != "" , ] 
myData.8h.sel <- myData.8h.sel [ !is.na(myData.8h.sel$DILIConcern) , ] 
myData.8h.sel <- myData.8h.sel [ myData.8h.sel$DILIConcern != "No_DILI_concern" , ] 
                                   
head(myData.8h.sel[ myData.8h.sel$DILIConcern == "oxidative stress", ])
unique(myData.8h.sel$DILIConcern)

myData.24h.sel <- myData.24h.sel [ myData.24h.sel$DILIConcern != "" , ] 
myData.24h.sel <- myData.24h.sel [ !is.na(myData.24h.sel$DILIConcern) , ] 
myData.24h.sel <- myData.24h.sel [ myData.24h.sel$DILIConcern != "No_DILI_concern" , ] 
unique(myData.24h.sel$DILIConcern)


myData.24h.sel[ myData.24h.sel$DILIConcern == "inflammation", "treatment"]
unique(myData.24h.sel$DILIConcern)
(myData.8h.sel$symbols)

# try a aheatmap package


require(reshape2)
head(myData.24h.sel)
# 24 hr data
myData.24h.sel$symbols_t <- paste(myData.24h.sel$symbols, myData.24h.sel$gene_type, sep ="_")


dataPlot24 <- myData.24h.sel[ , c("logFC", "treatment", "symbols_t" )]
dataPlot24$symbols_t <- gsub("nflammatory", "", dataPlot24$symbols_t)
dataPlot24$symbols_t <- gsub("xidative", "", dataPlot24$symbols_t) 

require(plyr)

annotation24 <- unique(myData.24h.sel[ , c("symbols_t", "gene_type", "treatment", "DILIConcern")])
annotation24$symbols_t <- gsub("nflammatory", "", annotation24$symbols_t)
annotation24$symbols_t <- gsub("xidative", "", annotation24$symbols_t)

dataPlot24.w <- dcast(dataPlot24,  treatment~ symbols_t, value.var= "logFC")

head(dataPlot24.w)
head(as.matrix(dataPlot24.w))
lapply(dataPlot24.w, class)
head(dataPlot24.w)
rownames(dataPlot24.w) <- dataPlot24.w$treatment
dataPlot24.w$treatment <- NULL

annotationGenes24 <- unique(annotation24[ , c("symbols_t", "gene_type")])
mt <- match( colnames(dataPlot24.w),annotationGenes24$symbols_t)
cbind(annotationGenes24$symbols_t[mt],colnames(dataPlot24.w))
annotationGenes24 <- annotationGenes24[, "gene_type" ][mt]
annotationGenes24 <- data.frame(gene_type = annotationGenes24)


#annotationGenes24 <- list(gene_type = annotationGenes24, genes = rownames())

#head(annotation24)
annotationComps24 <- unique(annotation24[ , c("treatment", "DILIConcern")])
annotationComps24$DILIConcern<-factor(annotationComps24$DILIConcern)
mt <- match( rownames(dataPlot24.w),annotationComps24$treatment)
cbind(as.character(annotationComps24$treatment[mt]),as.character(rownames(dataPlot24.w)))
annotationComps24 <- annotationComps24[, "DILIConcern" ][mt]
annotationComps24 <- data.frame(DILIConcern = annotationComps24)
annotationComps24$DILIConcern<-gsub("_", " ", annotationComps24$DILIConcern)






# 8 hr data
myData.8h.sel$symbols_t <- paste(myData.8h.sel$symbols, myData.8h.sel$gene_type, sep ="_")


dataPlot8 <- myData.8h.sel[ , c("logFC", "treatment", "symbols_t" )]
dataPlot8$symbols_t <- gsub("nflammatory", "", dataPlot8$symbols_t)
dataPlot8$symbols_t <- gsub("xidative", "", dataPlot8$symbols_t) 

require(plyr)

annotation8 <- unique(myData.8h.sel[ , c("symbols_t", "gene_type", "treatment", "DILIConcern")])
annotation8$symbols_t <- gsub("nflammatory", "", annotation8$symbols_t)
annotation8$symbols_t <- gsub("xidative", "", annotation8$symbols_t)

dataPlot8.w <- dcast(dataPlot8,  treatment~ symbols_t, value.var= "logFC")

head(dataPlot8.w)
head(as.matrix(dataPlot8.w))
lapply(dataPlot8.w, class)
head(dataPlot8.w)
rownames(dataPlot8.w) <- dataPlot8.w$treatment
dataPlot8.w$treatment <- NULL

annotationGenes8 <- unique(annotation8[ , c("symbols_t", "gene_type")])
mt <- match( colnames(dataPlot8.w),annotationGenes8$symbols_t)
cbind(annotationGenes8$symbols_t[mt],colnames(dataPlot8.w))
annotationGenes8 <- annotationGenes8[, "gene_type" ][mt]
annotationGenes8 <- data.frame(gene_type = annotationGenes8)


#annotationGenes24 <- list(gene_type = annotationGenes24, genes = rownames())

#head(annotation24)
annotationComps8 <- unique(annotation8[ , c("treatment", "DILIConcern")])
annotationComps8$DILIConcern<-factor(annotationComps8$DILIConcern)
mt <- match( rownames(dataPlot8.w),annotationComps8$treatment)
cbind(as.character(annotationComps8$treatment[mt]),as.character(rownames(dataPlot8.w)))
annotationComps8 <- annotationComps8[, "DILIConcern" ][mt]
annotationComps8 <- data.frame(DILIConcern = annotationComps8)
annotationComps8$DILIConcern<-gsub("_", " ", annotationComps8$DILIConcern)






#kmeans(x, centers, iter.max = 10, nstart = 1,
 #      algorithm = c("Hartigan-Wong", "Lloyd", "Forgy",
  #                   "MacQueen"), trace=FALSE)

# kmeans2 <- kmeans(dataPlot24.w, centers =9,  nstart =100, iter.max = 1000, algorithm = "Hartigan-Wong")
# kmeans2$cluster
# 
# 
# kmeansoutput <- cbind(kmeans2$cluster )
# o <- order(kmeans2$cluster)
# m2 <- (dataPlot24.w[o , ])
# #genes
# head(m2)
# kmeans2G <- kmeans(t(m2), center =4,  nstart =100, iter.max=1000, algorithm = "Hartigan-Wong")
# annotationComps24$DILIConcern<- as.character(annotationComps24$DILIConcern)
# oG <- order(kmeans2G$cluster)
# m2G <- (m2[ , oG ])
#=================================
#============== PLOT 24hrs
breaks <- (seq(min(dataPlot24.w), max(dataPlot24.w), length.out=101 )    )

# modify breaks: (visualize how)

plot(breaks)
lines(x=rep(50,5,101), y= seq(-10,10, length.out=101))
lines(x=rep(38,5,101), y= seq(-10,10, length.out=101), col="red")
lines(x=seq(0,101, length.out=101), y= rep(0,101) )
lines(x=seq(0,101, length.out=101), y= rep(1,101), col = "blue" )
lines(x=seq(0,101, length.out=101), y= rep(-1,101) , col= "blue")

#domL <- 2.5*sin(seq(0,0.75*pi, length.out=50))
#domR <- 3*sin(seq(-0.82*pi,0*pi, length.out=51))

##lines( -domL)
#lines(x=50:100,y= domR)
m.min <- min(dataPlot24.w)
m.max <-max(dataPlot24.w)
m.data <- data.frame(x = seq(1,101, length.out = 21), y =  c(m.min-0.1, m.min+ 2, m.min +3, m.min+4,  m.min+4.5,
                                                             m.min+4.7, -0.15, -0.05, 0.0, 0.05, 
                                                             0.1, 0.15, 0.2, 0.35, 0.45 ,0.7,
                                                             0.9, m.max-7, m.max-5, m.max-3, m.max+0.1))

m.data.sc <- data.frame(x = seq(1,101, length.out = 21), y =  c(m.min-0.1,-3.7, -2.6, -1.9,  -1.3,
                                                             -0.9, -0.55, -0.4, -0.3, -0.15, 
                                                             0.0, 0.15, 0.3, 0.4, 0.55 ,0.9,
                                                             1.2, 2, m.max-5, 5.5, m.max+0.1))

points(m.data.sc, col ="darkgreen")

points(m.data, col ="red")

x<- seq(1,101,length.out=101)

me.fit<-lm(y ~ I(x^5) + I(x^4)+I(x^3) +I(x^2) + I(x) , m.data.sc)

#?lm
#modBreaks <-breaks + c(-domL, domR )
#lines(modBreaks)
a<- me.fit$coefficients[2]
b<- me.fit$coefficients[3]
c<- me.fit$coefficients[4]
d<- me.fit$coefficients[5]
e<- me.fit$coefficients[6]
f<- me.fit$coefficients[1]


y <- function(x,a,b,c,d,e,f) {a*x^5 + b*x^4 + c*x^3 + d*x^2 + e*x + f}



modif<-y(x,a,b,c,d,e,f)
modif<-sort(modif)
lines(x,modif, col="red")
modBreaks <- modif #scaled
modBreaksRaw <- modif


m.min
m.max




DILIConcern <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
                "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#333333")

gene_type <- c("darkred","lightblue")
names(DILIConcern) <- c("fatal hepatotoxicity", "hyperbilirubinemia", "oxidative stress", "liver aminotransferase increase", "jaundice","cholestasis", "acute liver failure",  "inflammation", "liver necrosis")
names(gene_type) <- c("Inflammatory",  "Oxidative")

myColors <- list(DILIConcern = DILIConcern, gene_type=gene_type)




#aheatmap(as.matrix(t(m2G)), Rowv = NA, Colv=NA, annCol = annotationComps24, annRow = annotationGenes24,
#  scale = "row", breaks= breaks, annColors = myColors
#  )
#head(dataPlot24.w)
?aheatmap

head(dataPlot24.w)

'correlation', 'pearson', 'kendall', 'spearman', 'euclidean', 'maximum', 'manhattan', 'canberra', 'binary', 'minkowski'
'ward', 'single', 'complete', 'average', 'mcquitty', 'median', 'centroid'

#savePath <- "C:/Users/winks/Documents/own papers/NFKB Nrf2 Bram/figures/heatmaps both tp"
modBreaksRaw 

pdf(file = paste( "heatmap_24hrs_141_55_minkowskiWard_rowScaled_test2_none.pdf", sep ="/"))
aheatmap(as.matrix(t(dataPlot24.w)), Rowv = TRUE, Colv=TRUE, distfun = "minkowski",
         hclustfun = "ward" , annCol = annotationComps24, annRow = annotationGenes24,
  scale = "none", breaks= modBreaksRaw, annColors = myColors, fontsize = 1, width = 10, height=10, legend = TRUE,
  txt = round(as.matrix(t(dataPlot24.w)), digits=2)
  )
dev.off()
?scale
m.min
m.max



?aheatmap



###=========================================#
###======= PLOTTING 8hr=======================##
###=========================================#
###======= PLOTTING 8hr=======================##

breaks <- (seq(min(dataPlot8.w), max(dataPlot8.w), length.out=101 )    )

dataPlot8.w[1:5,1:5]
dataPlot24.w[1:5,1:5]

# modify breaks: (visualize how)
breaks[35]
plot(breaks)
lines(x=rep(50,5,101), y= seq(-10,10, length.out=101))
lines(x=rep(36,5,101), y= seq(-10,10, length.out=101), col="red")
lines(x=seq(0,101, length.out=101), y= rep(0,101) )
lines(x=seq(0,101, length.out=101), y= rep(1,101), col = "blue" )
lines(x=seq(0,101, length.out=101), y= rep(-1,101) , col= "blue")

#domL <- 2.5*sin(seq(0,0.75*pi, length.out=50))
#domR <- 3*sin(seq(-0.82*pi,0*pi, length.out=51))

##lines( -domL)
#lines(x=50:100,y= domR)
m.min <- min(dataPlot8.w)
m.max <-max(dataPlot8.w)
m.data <- data.frame(x = seq(1,101, length.out = 21), y =  c(m.min-0.1, m.min+ 2, -1, -0.35, -0.2 ,
                                                            -0.15 , -0.1, 0, 0.1, 0.15,
                                                             0.25, 0.35 , 0.5, 0.8, 1.2 ,1.6,
                                                             2.2, 3.5, m.max-2.5, m.max-1, m.max+0.1))
points(m.data, col ="red")

m.data.sc <- data.frame(x = seq(1,101, length.out = 21), y =  c(m.min-0.1,-3.7, -2.6, -1.9,  -1.3,
                                                             -0.9, -0.55, -0.4, -0.3, -0.15, 
                                                             0.0, 0.15, 0.3, 0.4, 0.55 ,0.9,
                                                             1.2, 2, m.max-5, 5.5, m.max+0.1))

points(m.data.sc, col ="darkgreen")

x<- seq(1,101,length.out=101)
me.fit<-lm(y ~ I(x^5) + I(x^4)+I(x^3) +I(x^2) + I(x) , m.data.sc)

#?lm
#modBreaks <-breaks + c(-domL, domR )
#lines(modBreaks)
a<- me.fit$coefficients[2]
b<- me.fit$coefficients[3]
c<- me.fit$coefficients[4]
d<- me.fit$coefficients[5]
e<- me.fit$coefficients[6]
f<- me.fit$coefficients[1]
y <- function(x,a,b,c,d,e,f) {a*x^5 + b*x^4 + c*x^3 + d*x^2 + e*x + f}



modif<-y(x,a,b,c,d,e,f)
modif<-sort(modif)
lines(x,modif, col="red")
modBreaks <- modif
modBreaksRaw <- modif
#kut eh
# nieuwe poging om "sneller" breaks te maken:
# 3 lineare stukken



DILIConcern <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
                "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#333333")

gene_type <- c("darkred","lightblue")
names(DILIConcern) <- c("fatal hepatotoxicity", "hyperbilirubinemia", "oxidative stress", "liver aminotransferase increase", "jaundice","cholestasis", "acute liver failure",  "inflammation", "liver necrosis")
names(gene_type) <- c("Inflammatory",  "Oxidative")

myColors <- list(DILIConcern = DILIConcern, gene_type=gene_type)


#aheatmap(as.matrix(t(m2G)), Rowv = NA, Colv=NA, annCol = annotationComps24, annRow = annotationGenes24,
#  scale = "row", breaks= breaks, annColors = myColors
#  )
#head(dataPlot24.w)
??aheatmap

head(dataPlot24.w)

'correlation', 'pearson', 'kendall', 'spearman', 'euclidean', 'maximum', 'manhattan', 'canberra', 'binary', 'minkowski'
'ward', 'single', 'complete', 'average', 'mcquitty', 'median', 'centroid'

#savePath <- "C:/Users/winks/Documents/own papers/NFKB Nrf2 Bram/figures/heatmaps both tp"
modBreaksRaw 

pdf(file = paste( "heatmap_8hrs_141_55_minkowskiWard_rowScaled.pdf", sep ="/"))
aheatmap(as.matrix(t(dataPlot8.w)), Rowv = TRUE, Colv=TRUE, distfun = "minkowski",
         hclustfun = "ward" , annCol = annotationComps24, annRow = annotationGenes24,
  scale = "row", breaks= modBreaks, annColors = myColors, fontsize = 8, width = 10, height=10, legend = FALSE
  )
dev.off()












========end bram paper



  # nu compounds orderen koloms gewijs naar dili-severity groepen. Dan uit elkaar trekken en kolommen clusteren. Dan weer samenvoegen en dan alleen de rijen clusteren.


unique(annotationComps24)
hist(1:100, col=cbbPalette$annColors)

?aheatmap

?aheatmap
length((dataPlot24.w$treatment))

head(dataPlot24)
airquality
aql <- melt(airquality, id.vars = c("Month", "Day"))

head(aql <- melt(airquality, id.vars = c("Month", "Day")))
head(aqw)
     aqw <- dcast(aql, Month + Day ~ variable)
aqw <- dcast(aql, month + day ~ variable)


# compound annotations
rownames(myAnot)<- myAnot$compounds
myAnot$compounds <-NULL
head(myAnot)
levels(myAnot$DILIConcern)
levels(myAnot$SeverityClass)
SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
DILIConcern <- c("grey","green", "orange", "darkred")
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
names(DILIConcern) <- c("unknown",  "No-DILI-Concern", "Less-DILI-Concern", "Most-DILI-Concern")
myColors <- list(SeverityClass = SeverityClass, DILIConcern=DILIConcern)

-log10(0.05)
max(myDF)
head(myAnot)
myVector<- cbind(as.character(myAnot$SeverityClass ), colnames(myDF)) 
myVectorInd<- myVector[,1] !="unknown"

myDFtest <- subset(myDF, select = myVectorInd )   #remove non-classified drugs
myDF <- myDFtest


class(myAnot)


myAnottest <- myAnot[ myVectorInd,]
myAnot <- myAnottest

head(myDFtest)
dim(myDF)
dim(myAnot)

#define colors for the pheatmap
## set color representation for specific values of the data distribution
quantile_range <- quantile(sqrt(as.matrix(myDF) ), probs = seq(0, 1, 0.1)) 
as.numeric(quantile_range)
#color_palette <- colorRampPalette(c(greenred(256)[1]  , greenred(256)[128], greenred(256)[256]  ))(length(quantile_range) - 1)

myDF <- sqrt(myDF)
head(myDF)
minV <- min(as.matrix(myDF))
maxV <- max(myDF)



hmcols<-colorRampPalette(c("white","blue"))(256)
white <- hmcols[1]
lightestBlue <-hmcols[30]
lightBlue <- hmcols[80]
blue <- hmcols[128]
darkBlue <- hmcols[156]
darkerBlue <- hmcols[180]
darkestBlue <-hmcols[256]

x<-1:10
hist(x, col = darkestBlue)

sqrt(1.33)
Breakwhite <- sqrt(1.29) 
BreaklightestBlue <- sqrt(2)
BreaklightBlue<- sqrt(2.5)
Breakblue <-  sqrt(3)
BreakdarkBlue <- sqrt( 3.5)
BreakdarkerBlue <- sqrt( 4)
BreakdArkestBlue <- maxV

max(myDF)

  myBreaks <-c(minV, Breakwhite, BreaklightestBlue, BreaklightBlue, Breakblue, BreakdarkBlue, BreakdarkerBlue, BreakdArkestBlue) 


color_palette <- c(white, lightestBlue, lightBlue, blue, darkBlue, darkerBlue, darkestBlue)

the distance measure to be used. This must be one of "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski". Any unambiguous substring can be given. and "correlation" - = pearson

the agglomeration method to be used. This should be (an unambiguous abbreviation of) one of "ward", "single", "complete", "average", "mcquitty", "median" or "centroid".

geneSet <-"IPA"
folderNameSampleSet <- "high_8hrs_selection"
#png(file = paste( geneSet, folderNameSampleSet ,zehDist, zehClust ,".png", sep = "_"), width = 5000, height = 10000, units = "px",  res = 8, type = "Xlib")




# = == this part is to divide myDF in two groups sDILI and nsDILI

head(myDF)
head(myAnot)

all(colnames(myDF )  == rownames(myAnot) )

indS <-  myAnot$DILIConcern %in% "Most-DILI-Concern"
myDF_sDILI <- myDF[, indS]
myAnot_sDILI <- myAnot[ indS, ]

indnS <-  myAnot$DILIConcern %in% "Less-DILI-Concern"
myDF_nsDILI <- myDF[ , indnS ]
myAnot_nsDILI <- myAnot[indnS , ]

head(myAnot_sDILI)

SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
myColors_sDILI <- list(SeverityClass = SeverityClass)



  head(myAnot_nsDILI)

SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
myColors_nsDILI <- list(SeverityClass = SeverityClass)

  
  
# this part is to store clustering of pheatmap into text file
?hclust
?dist
method  
zehDist <- c("manhattan","correlation", "euclidean")[1]
zehClust <- c("average", "ward", "complete", "centroid")[2]
dsDILI <- dist(myDF_sDILI , method = zehDist)
dnDILI <- dist(myDF_nsDILI , method = zehDist)
osDILI <- hclust(dsDILI, method = zehClust)
onsDILI <- hclust(dnDILI, method = zehClust)

TdsDILI <- dist(t(myDF_sDILI) , method = zehDist)
TdnDILI <- dist(t(myDF_nsDILI) , method = zehDist)
TosDILI <- hclust(TdsDILI, method = zehClust)
TonsDILI <- hclust(TdnDILI, method = zehClust)

ordered_myDF_sDILI <- myDF_sDILI[osDILI$order, TosDILI$order  ]
ordered_myDF_nsDILI <- myDF_nsDILI[onsDILI$order, TonsDILI$order  ]


# also try kmeans
kmeans2 <- kmeans(t(myDF_sDILI), center =3,  nstart =25)
kmeans2$cluster
o <- order(kmeans2$cluster)
m2 <- (t(myDF_sDILI)[o , ])
#genes
kmeans2G <- kmeans(myDF_sDILI, center =5,  nstart =25)
oG <- order(kmeans2G$cluster)
sDILIkmeans <- (m2[ , oG ])



kmeans2 <- kmeans(t(myDF_nsDILI), center =3,  nstart =25)
kmeans2$cluster
o <- order(kmeans2$cluster)
m2 <- (t(myDF_nsDILI)[o , ])
#genes
kmeans2G <- kmeans(myDF_nsDILI, center =3,  nstart =25)
oG <- order(kmeans2G$cluster)
nsDILIkmeans <- (m2[ , oG ])



pdf(file = paste("nsDILIkmeans_3_3", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
pheatmap(as.matrix(t(nsDILIkmeans)), col= color_palette, breaks =  myBreaks , main= paste(folderNameSampleSet,geneSet) ,  
         fontsize_row=2, fontsize_col =6, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
          legend = TRUE, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()


# now reorder rows to match each other and then recluster only the rows based on combined data

ind <- match( rownames(t(nsDILIkmeans) ), rownames(t(sDILIkmeans ) ) )
reordered_sDILIkmeans <- t(sDILIkmeans)[ind, ]
all(rownames(reordered_sDILIkmeans) == rownames(t(nsDILIkmeans) ) )
allData <- cbind(reordered_sDILIkmeans, t(nsDILIkmeans) )
head(allData)
allData<-as.data.frame(allData)


indBoth <-  myAnot$DILIConcern %in% c("Most-DILI-Concern", "Less-DILI-Concern")

myAnot_bothDILI <- myAnot[ indBoth, ]


SeverityClass <- c(  "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
DILIConcern <- c( "orange", "darkred")
names(SeverityClass) <- c(  "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
names(DILIConcern) <- c("Less-DILI-Concern", "Most-DILI-Concern")
myColors <- list(SeverityClass = SeverityClass, DILIConcern=DILIConcern)

zehDist <- c("manhattan","correlation", "euclidean")[1]
zehClust <- c("single" , "average", "ward", "complete", "centroid")[2]

pdf(file = paste("BothSepKmeans", geneSet, folderNameSampleSet ,zehDist, zehClust ,".pdf", sep = "_"), width = 8, height = 10, pointsize = 1)
fontsize_row = 3; fontsize_col = 3;
pheatmap(as.matrix(allData), col= color_palette, breaks =  myBreaks , main= paste(geneSet, folderNameSampleSet, sep ="_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = TRUE, cluster_cols = FALSE, clustering_distance_rows = zehDist, 
         clustering_method = zehClust,  
          legend = TRUE, fontsize = 4,
          legend_breaks = round(myBreaks, digits=2), 
         annotation = myAnot_bothDILI, annotation_colors = myColors,
         
         display_numbers = F )
dev.off()







#end of the splitting thing for Bob

zehDist <- c("manhattan","correlation", "euclidean")[3]
zehClust <- c("average", "ward", "complete", "centroid")[4]
pdf(file = paste( geneSet, folderNameSampleSet ,zehDist, zehClust ,"SEL.pdf", sep = "_"), width = 8, height = 10, pointsize = 1)
fontsize_row = 3; fontsize_col = 3;
pheatmap(as.matrix(myDF), col= color_palette, breaks =  myBreaks , main= paste(geneSet, folderNameSampleSet, sep ="_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = TRUE, cluster_cols = TRUE, clustering_distance_rows = zehDist, 
         clustering_distance_cols = zehDist, clustering_method = zehClust,  
          legend = TRUE, fontsize = 4,
          legend_breaks = round(myBreaks, digits=2), 
         annotation = myAnot, annotation_colors = myColors,
         
         display_numbers = F )
dev.off()

  ?pheatmap
#PCA plots: for complete code for pca plot including explaining variables (so genes) see : http://stackoverflow.com/questions/6578355/plotting-pca-biplot-with-ggplot2
?prcomp

myDFT<-t(myDF)

head(myDFT)
prin_comp<-rda(myDFT, scale=TRUE)
pca_scores<-scores(prin_comp)
head(pca_scores)

pdf(file = paste("PCA", folderNameSampleSet, ".pdf", sep  ="_"), height = 6, width = 6 )
plot(pca_scores$sites[,1],
     pca_scores$sites[,2],
     pch=21,
     bg=as.numeric(factor( myAnot$DILIConcern ) ),
     xlim=c(-2,2), 
     ylim=c(-2,2), 
     main = paste(geneSet,folderNameSampleSet) )
ordiellipse(prin_comp,myAnot$DILIConcern,conf=0.8)     
legend(x=1.3, y = 2, legend = c("unknown (blue)", "sDILI (red)", "nsDILI (black)", "nonDili (green)"  ), cex = 0.6)     
dev.off()

?legend
head(myAnot$DILIConcern)
as.numeric(myAnot$DILIConcern)
?plot  
    

To make ellipses, function ordiellipse() of package vegan is used. As arguments PCA analysis object and grouping variable must be provided. To control number of points included in ellipse, argument conf= can be used.

x<-1:10
hist(x, col = 2)


# clustering: how many clusters? take care if it is sample or gene based


#"euclidean" -2 , "maximum" -2, "manhattan" -5 , "canberra" -2, "binary" or "minkowski -2"
myDistDF <- dist(myDFT, method = "manhattan") # sample clustering
myDistDFM <- (as.matrix(myDistDF))
silEst <- silcheck(myDistDFM, diss = TRUE)
silEst

#"euclidean" -2 , "maximum" -2, "manhattan" -2 , "canberra" -2, "binary" or "minkowski -2"
myDistDF <- dist(myDF, method = "euclidean") # gene clustering
myDistDFM <- (as.matrix(myDistDF))
silEst <- silcheck(myDistDFM, diss = TRUE)
silEst

hobach.hobj <- hopach(myDFT, dmat = myDistDF)
hobach.hobj$clust$k 
hobach.hobj$cluster$sizes 

#conclusion: 2 or 4 gene clusters and 2,4,5 or 1 sample clusters

# partition based clustering: kmeans


#samples
kmeans2 <- kmeans(myDFT, center =3,  nstart =25)
kmeans2$cluster

kmeansoutput <- cbind(kmeans2$cluster )
o <- order(kmeans2$cluster)
m2 <- (myDFT[o , ])
#genes
kmeans2G <- kmeans(myDF, center =3,  nstart =25)

oG <- order(kmeans2G$cluster)
m2G <- (m2[ , oG ])


pdf(file = paste("kmeans_3_3", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
pheatmap(as.matrix(t(m2G)), col= color_palette, breaks =  myBreaks , main= paste(folderNameSampleSet,geneSet) ,  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
          legend = TRUE, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()




#8: fatal hepatoxicity  7: acute liver failure    6: liver necrosis    5: Jaundice	4: hyperbilirubinemia		3: liver aminotransferases increase		2: cholestasis	1: steatosis
    
    
    
write.table(kmeansoutput, file = paste(folderNameSampleSet, "kmeansC2", "csv", sep = "." ), sep = "," )



kmeans2 <- kmeans(myDFT, center =2,  nstart =25)
pdf(file = paste("Kmeans_clusterPlot", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
clusplot(myDFT, kmeans2$cluster, color= TRUE , shade=TRUE, 
         labels=2, lines=0, cex = 0.3, main = paste("clusterPlot", folderNameSampleSet, geneSet,  sep = "_") )
dev.off()

?clusplot
# partition based clustering: PAM

# euclidian and manhatten
pamK2.myDF <- pam(myDFT, diss = FALSE, metric= "manhattan", k = 2)
pamK2G.myDF <- pam(myDF, diss = FALSE, metric= "manhattan", k = 2)

o<-order(pamK2.myDF$cluster)
oG<-order(pamK2G.myDF$cluster)
pamM2 <- myDFT[o, oG  ]


pdf(file = paste("PAM_2_2_manh", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)

pheatmap(as.matrix(t(pamM2)), col= color_palette, breaks =  myBreaks , main= paste("PAM_S2_G3", folderNameSampleSet, geneSet,  sep = "_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE,  fontsize = 6,
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()




    pamResults <- cbind(pamK2.myDF$cluster)
write.table(pamResults, file = paste(folderNameSampleSet, "PAM_k5", "csv", sep = "." ), sep = "," )

pamK2.myDF <- pam(myDFT, diss = FALSE, metric= "euclidean", k = 2)
pdf(file = paste("PAMK2_clusterplot_eucl", folderNameSampleSet, geneSet, ",pdf", sep = "_") , width = 8, height = 8)
clusplot(myDFT, pamK2.myDF$cluster, color= TRUE , shade=TRUE, 
         labels=2, lines=0, cex = 0.3, main = paste("PAMK2_clusterPlot", folderNameSampleSet, geneSet,  sep = "_") )
dev.off()



# partition based clustering: SOM



SOM.myDF <- som(myDFT, grid = somgrid( 3, 3))
SOMG.myDF <- som(as.matrix(myDF), grid = somgrid( 2, 2 ))

table(SOM.myDF$unit.classif)

o<-order(SOM.myDF$unit.classif)
oG<-order(SOMG.myDF$unit.classif)
SOMM2 <- myDFT[o, oG  ]
head(t(SOMM2))


pdf(file = paste( "SOM_3_2", folderNameSampleSet, geneSet, ".pdf", sep = "_" ), width = 8, height = 10)

pheatmap(as.matrix(t(SOMM2)), col= color_palette, breaks =  myBreaks , main=paste( "SOM_S3_G4", folderNameSampleSet, geneSet, sep = "_" ),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE, display_numbers = F, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()

        cosangle
euclid
?distancematrix
myDist <- distancematrix(myDFT, d = "cosangle")
hobj <- hopach( myDFT, dmat = myDist)
myDist <- distancematrix(myDF, d = "cosangle")
hobjG <- hopach( myDF, dmat = myDist)

table(hobj$clustering$labels)
table(hobjG$clustering$labels)

o<-order(hobj$clustering$labels)
oG<-order(hobjG$clustering$labels)
HOP <- myDFT[o, oG  ]
head(t(HOP))


pdf(file = paste("HM_hopach_cosangle", folderNameSampleSet, geneSet,".pdf", sep = "_")  , width = 8, height = 10 )

pheatmap(as.matrix(t(HOP ) ), col= color_palette, breaks =  myBreaks , main= paste("HM_hopach_euclid", folderNameSampleSet, geneSet, sep = "_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE, display_numbers = F, fontsize = 6,
          legend_breaks = round(myBreaks, digits = 2), treeheight_row =0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()





  HOB1 <-cbind(rownames(myDFT[hobj$clust$labels == 300, ]), rep(1, sum(hobj$clust$labels == 300 )  ))
HOB2 <-cbind(rownames(myDFT[hobj$clust$labels == 250, ]), rep(2, sum(hobj$clust$labels == 250 )  ))

hopachgrps <- rbind(HOB1, HOB2)
write.table(hopachgrps, file = paste(folderNameSampleSet, "Hopach_2_biggest", "csv", sep = "." ), sep = "," )







```



