


DEG selection for ingenuity pathway analysis:
p.adj <0.05 with minimal of 500 genes.

1) compound level clustering: use p values, 0 for no pathway, -log10(p) > 1.33 and up for significant pathways
2) dili group level clustering: fraction of compounds where compound is in.

For both cases only use -log10(p) > 1.33  pathays lists

when loading the data make sure to obtain necessary info for both approaches
When myDF is completed enter other script for analysis & visualization





```{r }
rm(list=ls())


library(gdata)
library(stringr)


rm(list=ls())
rootPath <- "/home/winks/R_HOME/R_WORK/mArray BAC reporter studie/results/IPA/Metacore"
myFiles <- dir(rootPath)

myPaths <- paste(rootPath, myFiles, sep = "/")
ind.tox <- grepl("Liver", myFiles)
ind.all <- !ind.tox

myFiles.tox <-myPaths[ind.tox]
myFiles.all <- myPaths[ind.all]
tox.names = character()
my.list.tox = list()
for (i in 1: length(myFiles.tox)){

my.list.tox[[i]] <- read.xls(myFiles.tox[i], row.names = 1 )  
my.list.tox[[i]] <- my.list.tox[[i]][,1:2]  
my.list.tox[[i]] <- my.list.tox[[i]][my.list.tox[[i]][,'pValue'] <=0.05,]

my.name.tox <- str_match(myFiles.tox[i], '([^\\/]*)$'   )[1]
my.name.tox <- gsub(".xls", "", my.name.tox)
tox.names <- c(tox.names, my.name.tox)


}
all.names = character()
my.list.all = list()
for (i in 1: length(myFiles.all)){

my.list.all[[i]] <- read.xls(myFiles.all[i], row.names = 1 )  
my.list.all[[i]] <- my.list.all[[i]][,1:2]  
my.list.all[[i]] <- my.list.all[[i]][my.list.all[[i]][,'pValue'] <=0.05,]
my.name.all <- str_match(myFiles.all[i], '([^\\/]*)$'   )[1]
my.name.all <- gsub(".xls", "", my.name.all)
all.names <- c(all.names, my.name.all)
}


 
rawData.all <- do.call("rbind", my.list.all)
rawData.tox <- do.call("rbind", my.list.tox)
dim(rawData.tox)  
#1: first pull out all the unique pathways found in myList and create a data.frame
  #2: Then using matching to pull out appropriate values from myList

  
  

head(rawData.tox)

dim(rawData)
length(unique(rawData.all$Ingenuity_Canonical_Pathways))
# create empty data.frame with row names the unique pathways

myDF <- as.data.frame( matrix(NA, nrow = length( unique ( rawData.tox$Liver.toxicity.endpoint.processes )), ncol = length( myFiles.tox ) ) )
  rownames( myDF ) <- unique( rawData.tox$Liver.toxicity.endpoint.processes )
  
colnames( myDF ) <- tox.names
      head( myDF )
 


 for ( i in 1 : ncol( myDF )) {
   
   ind <- match( my.list.tox[[ i ]]$Liver.toxicity.endpoint.processes,  rownames( myDF ) ) 
   colnames(my.list.tox[[ i ]])[2]
   myDF[ind, i ] <- my.list.tox[[ i ]]$pValue 
 
   }
colnames(myDF) <- gsub(".Liver", "", colnames(myDF))

setwd("/home/winks/R_HOME/R_WORK/mArray BAC reporter studie/results/IPA")
write.table(file = "livermetacore_high_24hrs_subsetTest.csv", myDF, sep = ",", col.names = NA)

myDF[is.na( myDF )] <- 0
  head(myDF)

myDFsel <- subset(myDF, apply(as.matrix(myDF), MARGIN = 1, sum  ) > 50  ) #remove some pathways that are not doing much


head(myDFsel)[ , 1:2]
dim(myDFsel)
head(myDF)[, 1:2 ]
dim(myDF)

myDF <- myDFsel


```

bellow loading of compound - annotations - re-do after recreating myDF
```{r }
library(gdata)
options(stringsAsFactors=FALSE)


myDF <-read.delim("livermetacore_high_24hrs_subsetTest.csv", sep = ",", row.names=1, check.names=F)
myDF <- -log10(myDF)
colnames(myDF)

dim(myDF)
myDF[is.na( myDF )] <- 0
  head(myDF)

myDFsel <- subset(myDF, apply(as.matrix(myDF), MARGIN = 1, sum  ) > 30  ) #remove some pathways that are not doing much
dim(myDFsel)
myDF <- myDFsel
head(myDFsel)


myGroups <- read.xls( file.path( "/home/winks/compounds for screen",  "TGPDrugsAnnotatedbyLabel_ForSteven.xls"), sep =",", header = T)

myAnot <- data.frame(compounds = colnames(myDF) ) #currently used compounds

myGroups <- subset( myGroups, select = c("Name_TGP","SeverityClass", "DILIConcern" ))
head(myGroups)
colnames(myGroups)<- c("compound", "SeverityClass", "DILIConcern" )   

unique(myGroups$DILIConcern)
myGroups$DILIConcern <- gsub("Less-DILI-concern", "Less-DILI-Concern",   myGroups$DILIConcern )
myGroups$DILIConcern <- gsub("Most-DILI-concern", "Most-DILI-Concern",  myGroups$DILIConcern )
myGroups$DILIConcern[myGroups$DILIConcern == ""] <-"unknown"
#8: fatal hepatoxicity  7: acute liver failure    6: liver necrosis    5: Jaundice	4: hyperbilirubinemia		3: liver aminotransferases increase		2: cholestasis	1: steatosis

myGroups$SeverityClass[myGroups$SeverityClass == ""] <-"unknown"
myGroups$SeverityClass <-gsub("0", "No_DILI_concern",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("1", "steatosis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("2", "cholestasis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("3", "liver_aminotransferase_increase",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("4", "hyperbilirubinemia",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("5", "jaundice",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("6", "liver_necrosis",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("7", "acute_liver_failure",myGroups$SeverityClass   )
myGroups$SeverityClass <-gsub("8", "fatal_hepatotoxicity",myGroups$SeverityClass   )

# parsing problem for TNF
myGroups$compound[ 167 ] <- "TNF"
myGroups<-myGroups[-171,]
sort(myGroups$compound)

ind<-match(myAnot[,1], myGroups$compound)


if (length(myAnot[,1]) != sum(!is.na(ind))) 
  {
  stop("not guud, some compound did not match in the two lists!")
  }

myAnot$SeverityClass <- myGroups$SeverityClass[ ind ]
myAnot$DILIConcern <- myGroups$DILIConcern[ ind ]

myAnot$SeverityClass <- factor(myAnot$SeverityClass)
myAnot$DILIConcern <- factor(myAnot$DILIConcern)
myAnot$compounds <- factor(myAnot$compounds)


```


```{r}

library(pheatmap)   
library(gplots)
library("Biobase")
library("RColorBrewer")
library(cluster)
library(fpc)
library(ggplot2)
library(kohonen)
library(hopach)
library(vegan)


head(myAnot)
head(myDF)
dim(myDF)

# compound annotations
rownames(myAnot)<- myAnot$compounds
myAnot$compounds <-NULL
head(myAnot)
levels(myAnot$DILIConcern)
levels(myAnot$SeverityClass)
SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
DILIConcern <- c("grey","green", "orange", "darkred")
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
names(DILIConcern) <- c("unknown",  "No-DILI-Concern", "Less-DILI-Concern", "Most-DILI-Concern")
myColors <- list(SeverityClass = SeverityClass, DILIConcern=DILIConcern)

-log10(0.05)
max(myDF)
head(myAnot)
myVector<- cbind(as.character(myAnot$SeverityClass ), colnames(myDF)) 
myVectorInd<- myVector[,1] !="unknown"
myVectorInd[is.na(myVectorInd)]<-FALSE
myDFtest <- subset(myDF, select = myVectorInd )   #remove non-classified drugs

myDF <- myDFtest


class(myAnot)


myAnottest <- myAnot[ myVectorInd,]
myAnot <- myAnottest

head(myDFtest)
head(myDF)
head(myAnot)

#define colors for the pheatmap
## set color representation for specific values of the data distribution
quantile_range <- quantile(sqrt(as.matrix(myDF) ), probs = seq(0, 1, 0.1)) 
as.numeric(quantile_range)
#color_palette <- colorRampPalette(c(greenred(256)[1]  , greenred(256)[128], greenred(256)[256]  ))(length(quantile_range) - 1)

myDF <- sqrt(myDF)
head(myDF)
minV <- min(as.matrix(myDF))
maxV <- max(myDF)



hmcols<-colorRampPalette(c("white","blue"))(256)


white <- hmcols[1]
lightestBlue <-hmcols[30]
lightBlue <- hmcols[80]
blue <- hmcols[128]
darkBlue <- hmcols[156]
darkerBlue <- hmcols[180]
darkestBlue <-hmcols[256]

x<-1:10
hist(x, col = darkestBlue)

sqrt(1.33)
Breakwhite <- sqrt(1.29) 
BreaklightestBlue <- sqrt(2)
BreaklightBlue<- sqrt(2.5)
Breakblue <-  sqrt(3)
BreakdarkBlue <- sqrt( 3.5)
BreakdarkerBlue <- sqrt( 4)
BreakdArkestBlue <- maxV

max(myDF)

  myBreaks <-c(minV, Breakwhite, BreaklightestBlue, BreaklightBlue, Breakblue, BreakdarkBlue, BreakdarkerBlue, BreakdArkestBlue) 


color_palette <- c(white, lightestBlue, lightBlue, blue, darkBlue, darkerBlue, darkestBlue)

the distance measure to be used. This must be one of "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski". Any unambiguous substring can be given. and "correlation" - = pearson

the agglomeration method to be used. This should be (an unambiguous abbreviation of) one of "ward", "single", "complete", "average", "mcquitty", "median" or "centroid".

geneSet <-"IPA"
folderNameSampleSet <- "all_MC_high_24hrs_selection"
#png(file = paste( geneSet, folderNameSampleSet ,zehDist, zehClust ,".png", sep = "_"), width = 5000, height = 10000, units = "px",  res = 8, type = "Xlib")




# = == this part is to divide myDF in two groups sDILI and nsDILI

head(myDF)
head(myAnot)

all(colnames(myDF )  == rownames(myAnot) )

indS <-  myAnot$DILIConcern %in% "Most-DILI-Concern"
myDF_sDILI <- myDF[, indS]
myAnot_sDILI <- myAnot[ indS, ]

indnS <-  myAnot$DILIConcern %in% "Less-DILI-Concern"
myDF_nsDILI <- myDF[ , indnS ]
myAnot_nsDILI <- myAnot[indnS , ]

head(myAnot_sDILI)

SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
myColors_sDILI <- list(SeverityClass = SeverityClass)



  head(myAnot_nsDILI)

SeverityClass <- c("grey",  "green", "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
names(SeverityClass) <- c("unknown", "No_DILI_concern", "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
myColors_nsDILI <- list(SeverityClass = SeverityClass)

  
  
# this part is to store clustering of pheatmap into text file
?hclust
?dist
method  
zehDist <- c("manhattan","correlation", "euclidean")[1]
zehClust <- c("average", "ward", "complete", "centroid")[2]
dsDILI <- dist(myDF_sDILI , method = zehDist)
dnDILI <- dist(myDF_nsDILI , method = zehDist)
osDILI <- hclust(dsDILI, method = zehClust)
onsDILI <- hclust(dnDILI, method = zehClust)

TdsDILI <- dist(t(myDF_sDILI) , method = zehDist)
TdnDILI <- dist(t(myDF_nsDILI) , method = zehDist)
TosDILI <- hclust(TdsDILI, method = zehClust)
TonsDILI <- hclust(TdnDILI, method = zehClust)

ordered_myDF_sDILI <- myDF_sDILI[osDILI$order, TosDILI$order  ]
ordered_myDF_nsDILI <- myDF_nsDILI[onsDILI$order, TonsDILI$order  ]


# also try kmeans
kmeans2 <- kmeans(t(myDF_sDILI), center =3,  nstart =25)
kmeans2$cluster
o <- order(kmeans2$cluster)
m2 <- (t(myDF_sDILI)[o , ])
#genes
kmeans2G <- kmeans(myDF_sDILI, center =5,  nstart =25)
oG <- order(kmeans2G$cluster)
sDILIkmeans <- (m2[ , oG ])



kmeans2 <- kmeans(t(myDF_nsDILI), center =3,  nstart =25)
kmeans2$cluster
o <- order(kmeans2$cluster)
m2 <- (t(myDF_nsDILI)[o , ])
#genes
kmeans2G <- kmeans(myDF_nsDILI, center =3,  nstart =25)
oG <- order(kmeans2G$cluster)
nsDILIkmeans <- (m2[ , oG ])



pdf(file = paste("nsDILIkmeans_3_3", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
pheatmap(as.matrix(t(nsDILIkmeans)), col= color_palette, breaks =  myBreaks , main= paste(folderNameSampleSet,geneSet) ,  
         fontsize_row=2, fontsize_col =6, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
          legend = TRUE, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()


# now reorder rows to match each other and then recluster only the rows based on combined data

ind <- match( rownames(t(nsDILIkmeans) ), rownames(t(sDILIkmeans ) ) )
reordered_sDILIkmeans <- t(sDILIkmeans)[ind, ]
all(rownames(reordered_sDILIkmeans) == rownames(t(nsDILIkmeans) ) )
allData <- cbind(reordered_sDILIkmeans, t(nsDILIkmeans) )
head(allData)
allData<-as.data.frame(allData)


indBoth <-  myAnot$DILIConcern %in% c("Most-DILI-Concern", "Less-DILI-Concern")

myAnot_bothDILI <- myAnot[ indBoth, ]


SeverityClass <- c(  "yellow", "orange", "pink","red",  "lightblue" ,"blue", "purple","black" )
DILIConcern <- c( "orange", "darkred")
names(SeverityClass) <- c(  "cholestasis", "liver_aminotransferase_increase", "hyperbilirubinemia",
                         "jaundice", "liver_necrosis",  "acute_liver_failure", "fatal_hepatotoxicity", "WD")
names(DILIConcern) <- c("Less-DILI-Concern", "Most-DILI-Concern")
myColors <- list(SeverityClass = SeverityClass, DILIConcern=DILIConcern)

zehDist <- c("manhattan","correlation", "euclidean")[1]
zehClust <- c("single" , "average", "ward", "complete", "centroid")[2]

pdf(file = paste("BothSepKmeans", geneSet, folderNameSampleSet ,zehDist, zehClust ,".pdf", sep = "_"), width = 8, height = 10, pointsize = 1)
fontsize_row = 3; fontsize_col = 3;
pheatmap(as.matrix(allData), col= color_palette, breaks =  myBreaks , main= paste(geneSet, folderNameSampleSet, sep ="_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = TRUE, cluster_cols = FALSE, clustering_distance_rows = zehDist, 
         clustering_method = zehClust,  
          legend = TRUE, fontsize = 4,
          legend_breaks = round(myBreaks, digits=2), 
         annotation = myAnot_bothDILI, annotation_colors = myColors,
         
         display_numbers = F )
dev.off()







#end of the splitting thing for Bob

zehDist <- c("manhattan","correlation", "euclidean")[3]
zehClust <- c("average", "ward", "complete", "centroid")[4]
pdf(file = paste( geneSet, folderNameSampleSet ,zehDist, zehClust ,"SEL.pdf", sep = "_"), width = 8, height = 10, pointsize = 1)
fontsize_row = 3; fontsize_col = 3;
pheatmap(as.matrix(myDF), col= color_palette, breaks =  myBreaks , main= paste(geneSet, folderNameSampleSet, sep ="_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = TRUE, cluster_cols = TRUE, clustering_distance_rows = zehDist, 
         clustering_distance_cols = zehDist, clustering_method = zehClust,  
          legend = TRUE, fontsize = 4,
          legend_breaks = round(myBreaks, digits=2), 
         annotation = myAnot, annotation_colors = myColors,
         
         display_numbers = F )
dev.off()

  ?pheatmap
#PCA plots: for complete code for pca plot including explaining variables (so genes) see : http://stackoverflow.com/questions/6578355/plotting-pca-biplot-with-ggplot2
?prcomp

myDFT<-t(myDF)

head(myDFT)
prin_comp<-rda(myDFT, scale=TRUE)
pca_scores<-scores(prin_comp)
head(pca_scores)

pdf(file = paste("PCA", folderNameSampleSet, ".pdf", sep  ="_"), height = 6, width = 6 )
plot(pca_scores$sites[,1],
     pca_scores$sites[,2],
     pch=21,
     bg=as.numeric(factor( myAnot$DILIConcern ) ),
     xlim=c(-2,2), 
     ylim=c(-2,2), 
     main = paste(geneSet,folderNameSampleSet) )
ordiellipse(prin_comp,myAnot$DILIConcern,conf=0.8)     
legend(x=1.3, y = 2, legend = c("unknown (blue)", "sDILI (red)", "nsDILI (black)", "nonDili (green)"  ), cex = 0.6)     
dev.off()

?legend
head(myAnot$DILIConcern)
as.numeric(myAnot$DILIConcern)
?plot  
    

To make ellipses, function ordiellipse() of package vegan is used. As arguments PCA analysis object and grouping variable must be provided. To control number of points included in ellipse, argument conf= can be used.

x<-1:10
hist(x, col = 2)


# clustering: how many clusters? take care if it is sample or gene based


#"euclidean" -2 , "maximum" -2, "manhattan" -5 , "canberra" -2, "binary" or "minkowski -2"
myDistDF <- dist(myDFT, method = "manhattan") # sample clustering
myDistDFM <- (as.matrix(myDistDF))
silEst <- silcheck(myDistDFM, diss = TRUE)
silEst

#"euclidean" -2 , "maximum" -2, "manhattan" -2 , "canberra" -2, "binary" or "minkowski -2"
myDistDF <- dist(myDF, method = "euclidean") # gene clustering
myDistDFM <- (as.matrix(myDistDF))
silEst <- silcheck(myDistDFM, diss = TRUE)
silEst

hobach.hobj <- hopach(myDFT, dmat = myDistDF)
hobach.hobj$clust$k 
hobach.hobj$cluster$sizes 

#conclusion: 2 or 4 gene clusters and 2,4,5 or 1 sample clusters

# partition based clustering: kmeans


#samples
kmeans2 <- kmeans(myDFT, center =3,  nstart =25)
kmeans2$cluster

kmeansoutput <- cbind(kmeans2$cluster )
o <- order(kmeans2$cluster)
m2 <- (myDFT[o , ])
#genes
kmeans2G <- kmeans(myDF, center =3,  nstart =25)

oG <- order(kmeans2G$cluster)
m2G <- (m2[ , oG ])


pdf(file = paste("kmeans_3_3", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
pheatmap(as.matrix(t(m2G)), col= color_palette, breaks =  myBreaks , main= paste(folderNameSampleSet,geneSet) ,  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
          legend = TRUE, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()




#8: fatal hepatoxicity  7: acute liver failure    6: liver necrosis    5: Jaundice	4: hyperbilirubinemia		3: liver aminotransferases increase		2: cholestasis	1: steatosis
    
    
    
write.table(kmeansoutput, file = paste(folderNameSampleSet, "kmeansC2", "csv", sep = "." ), sep = "," )



kmeans2 <- kmeans(myDFT, center =2,  nstart =25)
pdf(file = paste("Kmeans_clusterPlot", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)
clusplot(myDFT, kmeans2$cluster, color= TRUE , shade=TRUE, 
         labels=2, lines=0, cex = 0.3, main = paste("clusterPlot", folderNameSampleSet, geneSet,  sep = "_") )
dev.off()

?clusplot
# partition based clustering: PAM

# euclidian and manhatten
pamK2.myDF <- pam(myDFT, diss = FALSE, metric= "manhattan", k = 2)
pamK2G.myDF <- pam(myDF, diss = FALSE, metric= "manhattan", k = 2)

o<-order(pamK2.myDF$cluster)
oG<-order(pamK2G.myDF$cluster)
pamM2 <- myDFT[o, oG  ]


pdf(file = paste("PAM_2_2_manh", folderNameSampleSet, geneSet, ".pdf", sep = "_") , width = 8, height = 8)

pheatmap(as.matrix(t(pamM2)), col= color_palette, breaks =  myBreaks , main= paste("PAM_S2_G3", folderNameSampleSet, geneSet,  sep = "_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE,  fontsize = 6,
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()




    pamResults <- cbind(pamK2.myDF$cluster)
write.table(pamResults, file = paste(folderNameSampleSet, "PAM_k5", "csv", sep = "." ), sep = "," )

pamK2.myDF <- pam(myDFT, diss = FALSE, metric= "euclidean", k = 2)
pdf(file = paste("PAMK2_clusterplot_eucl", folderNameSampleSet, geneSet, ",pdf", sep = "_") , width = 8, height = 8)
clusplot(myDFT, pamK2.myDF$cluster, color= TRUE , shade=TRUE, 
         labels=2, lines=0, cex = 0.3, main = paste("PAMK2_clusterPlot", folderNameSampleSet, geneSet,  sep = "_") )
dev.off()



# partition based clustering: SOM



SOM.myDF <- som(myDFT, grid = somgrid( 3, 3))
SOMG.myDF <- som(as.matrix(myDF), grid = somgrid( 2, 2 ))

table(SOM.myDF$unit.classif)

o<-order(SOM.myDF$unit.classif)
oG<-order(SOMG.myDF$unit.classif)
SOMM2 <- myDFT[o, oG  ]
head(t(SOMM2))


pdf(file = paste( "SOM_3_2", folderNameSampleSet, geneSet, ".pdf", sep = "_" ), width = 8, height = 10)

pheatmap(as.matrix(t(SOMM2)), col= color_palette, breaks =  myBreaks , main=paste( "SOM_S3_G4", folderNameSampleSet, geneSet, sep = "_" ),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE, display_numbers = F, fontsize = 6, 
          legend_breaks = round(myBreaks, digits = 2), treeheight_row = 0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()

        cosangle
euclid
?distancematrix
myDist <- distancematrix(myDFT, d = "cosangle")
hobj <- hopach( myDFT, dmat = myDist)
myDist <- distancematrix(myDF, d = "cosangle")
hobjG <- hopach( myDF, dmat = myDist)

table(hobj$clustering$labels)
table(hobjG$clustering$labels)

o<-order(hobj$clustering$labels)
oG<-order(hobjG$clustering$labels)
HOP <- myDFT[o, oG  ]
head(t(HOP))


pdf(file = paste("HM_hopach_cosangle", folderNameSampleSet, geneSet,".pdf", sep = "_")  , width = 8, height = 10 )

pheatmap(as.matrix(t(HOP ) ), col= color_palette, breaks =  myBreaks , main= paste("HM_hopach_euclid", folderNameSampleSet, geneSet, sep = "_"),  
         fontsize_row=fontsize_row, fontsize_col =fontsize_col, border_color=NA, scale = "none",
         cluster_rows = FALSE, cluster_cols = FALSE,  
         
          legend = TRUE, display_numbers = F, fontsize = 6,
          legend_breaks = round(myBreaks, digits = 2), treeheight_row =0, treeheight_col = 0,
         annotation = myAnot, annotation_colors = myColors
         )

dev.off()





  HOB1 <-cbind(rownames(myDFT[hobj$clust$labels == 300, ]), rep(1, sum(hobj$clust$labels == 300 )  ))
HOB2 <-cbind(rownames(myDFT[hobj$clust$labels == 250, ]), rep(2, sum(hobj$clust$labels == 250 )  ))

hopachgrps <- rbind(HOB1, HOB2)
write.table(hopachgrps, file = paste(folderNameSampleSet, "Hopach_2_biggest", "csv", sep = "." ), sep = "," )







```


